\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{microtype}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue
}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{conjecture}[theorem]{Conjecture}

\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\supp}{\mathrm{supp}}
\newcommand{\softmax}{\mathrm{softmax}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\Pone}{\mathcal{P}_1}

\title{Theory Skeleton: Mean-Field Training and Gauge Freedom\\in Gaussian Mixture Density Networks with Natural Parameters}
\author{Working draft}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This note extends the softplus natural-parameter program from a single Gaussian
head to a fixed-$K$ Gaussian mixture density network (MDN). Each component is
parameterized by Gaussian natural parameters
$(\eta_{1,k},\eta_{2,k})=(\mu_k/\sigma_k^2,-1/(2\sigma_k^2))$, while the second
coordinate is implemented in practice by a raw scale field passed through a
softplus map so that $\eta_{2,k}<0$ automatically. The first theorem proved
here is a finite-horizon shared-path result for plain SGD: in the joint
wide/small-step limit, independently initialized wide MDNs track the same
deterministic labeled measure path $\rho_t$ on every fixed interval $[0,T]$.
The second theorem explains why this still does \emph{not} solve the
model-synthesis problem for plain MDNs: even at identical predictive
distributions, no global permutation of the components may align two mixture
representations, and already at the output-head level a linear interpolation can
incur a strictly positive loss barrier. The correct interpretation is that
deterministic mean-field dynamics suppress run-to-run gauge randomness on fixed
finite horizons, but do not canonically remove mixture non-identifiability.
This sharply separates the $K=1$ and $K>1$ stories: the training-side
mean-field bridge survives, but the synthesis-side canonical-alignment premise
used in the single-Gaussian argument can already fail at $K=2$.
\end{abstract}

\section{Executive Summary}

The MDN extension should be read as two separate messages.

\begin{enumerate}[label=(\roman*)]
  \item \textbf{Training-side message.} For a fixed number of mixture
  components $K$, a two-layer Gaussian MDN with softmax gates, Gaussian natural
  parameters, and softplus raw scale heads still admits a finite-horizon
  mean-field program in the raw particle variables. In the plain-SGD
  scaling used in two-layer LMC work, this program can be closed with the same
  training-side structure as in the single-Gaussian case.
  \item \textbf{Synthesis-side message.} The deterministic mean-field path is a
  labeled path. It gives repeatability for a chosen initialization law, but it
  does not produce a canonical component decomposition. Global-permutation model
  synthesis can therefore fail even between predictors that represent exactly
  the same conditional density. This is the first place where the $K=1$ and
  $K>1$ stories diverge.
\end{enumerate}

The intended downstream use is clear. The first message supplies the same kind
of training-side bridge that LMC arguments need in the two-layer scalar case
\cite{FerbachEtAl2024}. The second message explains why plain MDNs remain
harder: finite-width randomness can select different representatives of the same
predictive mixture class, and global permutation is too weak to compare them in
general.

\section{Gaussian MDN with Natural Parameters}

Let $(X,Y)\sim P$ with $X\in\mc{X}\subseteq \R^d$ and $Y\in\R$. Fix a component
count $K\ge 2$. We consider the standard fixed-$K$ Gaussian MDN architecture in
the sense of Bishop~\cite{Bishop2006}, but parameterized directly by Gaussian
natural coordinates. Define
\[
  A_{\rm G}(\eta_1,\eta_2)
  :=
  -\frac{\eta_1^2}{4\eta_2}
  + \frac12 \log\!\Big(\frac{-\pi}{\eta_2}\Big),
  \qquad \eta_2<0,
\]
and write the component negative log-likelihood as
\[
  \ell_{\rm G}(\eta;y)
  :=
  A_{\rm G}(\eta)-\eta_1 y-\eta_2 y^2.
\]

Let $\varphi:\R^p\times \mc{X}\to \R$ be a scalar hidden feature map. A single particle
is
\[
  \theta=(\beta,a,c,\xi)\in \Theta:=\R^K\times\R^K\times\R^K\times \R^p,
\]
where $\beta=(\beta_1,\dots,\beta_K)$ controls the gate logits,
$a=(a_1,\dots,a_K)$ the first natural coordinates, and
$c=(c_1,\dots,c_K)$ the raw scale coordinates.

Fix $\lambda_{\min}>0$ and
\[
  \psi_{\rm sp}(s):=\lambda_{\min}+\log(1+e^s).
\]
At finite width $N$, define for each component $k$
\begin{align}
  g_{k,N}(x)
  &:=
  \frac1N\sum_{i=1}^N \beta_{i,k}\,\varphi(\xi_i,x),
  \\
  \eta_{1,k,N}(x)
  &:=
  \frac1N\sum_{i=1}^N a_{i,k}\,\varphi(\xi_i,x),
  \\
  s_{k,N}(x)
  &:=
  \frac1N\sum_{i=1}^N c_{i,k}\,\varphi(\xi_i,x),
  \\
  \eta_{2,k,N}(x)
  &:=
  -\psi_{\rm sp}(s_{k,N}(x)).
\end{align}
The gate weights are
\[
  \pi_{k,N}(x)
  :=
  \frac{e^{g_{k,N}(x)}}{\sum_{\ell=1}^K e^{g_{\ell,N}(x)}},
\]
and the predictive conditional density is
\begin{equation}
  p_N(y|x)
  :=
  \sum_{k=1}^K
  \pi_{k,N}(x)\,
  \exp\!\big(
    \eta_{1,k,N}(x)y + \eta_{2,k,N}(x)y^2 - A_{\rm G}(\eta_{k,N}(x))
  \big).
  \label{eq:finite-mdn-density}
\end{equation}

The corresponding mean-field fields for a probability measure
$\rho\in \Pone(\Theta)$ are
\begin{align}
  g_{k,\rho}(x)
  &:=
  \int_\Theta \beta_k \varphi(\xi,x)\,\rho(d\theta),
  \\
  \eta_{1,k,\rho}(x)
  &:=
  \int_\Theta a_k \varphi(\xi,x)\,\rho(d\theta),
  \\
  s_{k,\rho}(x)
  &:=
  \int_\Theta c_k \varphi(\xi,x)\,\rho(d\theta),
  \\
  \eta_{2,k,\rho}(x)
  &:=
  -\psi_{\rm sp}(s_{k,\rho}(x)).
\end{align}
We abbreviate $\eta_{k,\rho}(x):=(\eta_{1,k,\rho}(x),\eta_{2,k,\rho}(x))$ and
\[
  \pi_{k,\rho}(x)
  :=
  \frac{e^{g_{k,\rho}(x)}}{\sum_{\ell=1}^K e^{g_{\ell,\rho}(x)}}.
\]

\begin{proposition}[Automatic admissibility of the second natural coordinates]
\label{prop:mdn-domain}
For every probability measure $\rho\in \Pone(\Theta)$, every finite-width
realization, every component $k$, and every $x\in \mc{X}$,
\[
  \eta_{2,k,\rho}(x)\le -\lambda_{\min}<0,
  \qquad
  \eta_{2,k,N}(x)\le -\lambda_{\min}<0.
\]
\end{proposition}

\begin{proof}
By definition, $\psi_{\rm sp}(s)\ge \lambda_{\min}$ for every $s\in\R$.
\end{proof}

For later use define the Gaussian component derivatives
\begin{align}
  q_1(\eta;y)
  &:=
  \partial_{\eta_1}\ell_{\rm G}(\eta;y)
  =
  -\frac{\eta_1}{2\eta_2}-y,
  \\
  q_2(\eta;y)
  &:=
  \partial_{\eta_2}\ell_{\rm G}(\eta;y)
  =
  \frac{\eta_1^2}{4\eta_2^2}
  -
  \frac{1}{2\eta_2}
  -
  y^2.
\end{align}

\begin{definition}[Responsibilities]
For a measure $\rho\in \Pone(\Theta)$, define the responsibility of component
$k$ by
\begin{equation}
  r_{k,\rho}(x,y)
  :=
  \frac{
    \pi_{k,\rho}(x)\,
    \exp\!\big(\eta_{1,k,\rho}(x)y+\eta_{2,k,\rho}(x)y^2
    -A_{\rm G}(\eta_{k,\rho}(x))\big)
  }{
    \sum_{\ell=1}^K
    \pi_{\ell,\rho}(x)\,
    \exp\!\big(\eta_{1,\ell,\rho}(x)y+\eta_{2,\ell,\rho}(x)y^2
    -A_{\rm G}(\eta_{\ell,\rho}(x))\big)
  }.
  \label{eq:responsibility}
\end{equation}
\end{definition}

\begin{lemma}[MDN differential identities]
\label{lem:mdn-differential-identities}
Let
\[
  \ell_{\rm MDN}(g,\eta;y)
  :=
  -\log\!\left[
    \sum_{k=1}^K
    \pi_k(g)\,
    \exp\!\big(
      \eta_{1,k}y+\eta_{2,k}y^2-A_{\rm G}(\eta_k)
    \big)
  \right],
\]
where $\pi_k(g)=e^{g_k}/\sum_{\ell=1}^K e^{g_\ell}$. Then
\begin{align}
  \partial_{g_k}\ell_{\rm MDN}(g,\eta;y)
  &=
  \pi_k(g)-r_k(g,\eta;y),
  \label{eq:mdn-gate-derivative}
  \\
  \partial_{\eta_{1,k}}\ell_{\rm MDN}(g,\eta;y)
  &=
  r_k(g,\eta;y)\,q_1(\eta_k;y),
  \label{eq:mdn-eta1-derivative}
  \\
  \partial_{\eta_{2,k}}\ell_{\rm MDN}(g,\eta;y)
  &=
  r_k(g,\eta;y)\,q_2(\eta_k;y),
  \label{eq:mdn-eta2-derivative}
\end{align}
where $r_k(g,\eta;y)$ denotes the corresponding responsibility.
\end{lemma}

\begin{proof}
Write
\[
  Z(g,\eta;y)
  :=
  \sum_{\ell=1}^K
  \exp\!\big(
    g_\ell-\log\!\sum_{j=1}^K e^{g_j}
    +\eta_{1,\ell}y+\eta_{2,\ell}y^2-A_{\rm G}(\eta_\ell)
  \big).
\]
Then $\ell_{\rm MDN}=-\log Z$. Differentiating $-\log Z$ with respect to $g_k$
gives
\[
  -\frac{1}{Z}\partial_{g_k}Z
  =
  -\frac{1}{Z}\sum_{\ell=1}^K
  \exp(\cdots)(\delta_{k\ell}-\pi_k)
  =
  \pi_k-r_k.
\]
Differentiating with respect to $\eta_{1,k}$ and $\eta_{2,k}$ gives
\[
  -\frac{1}{Z}\partial_{\eta_{j,k}}Z
  =
  r_k\,\partial_{\eta_j}\ell_{\rm G}(\eta_k;y),
  \qquad j\in\{1,2\},
\]
which is exactly \eqref{eq:mdn-eta1-derivative} and
\eqref{eq:mdn-eta2-derivative}.
\end{proof}

\section{Training-Side Theorem in a Labeled Gauge}

The practical training variables are the raw particle parameters
$(\beta_i,a_i,c_i,\xi_i)$, not the transformed natural coordinates. The first
task is therefore not to quotient the mixture symmetry, but to verify that the
raw MDN drift fits the same finite-horizon SGD mean-field framework used in the
single-Gaussian case.

\begin{assumption}[Finite-horizon MDN SGD regime]
\label{ass:mdn-sgd}
Fix a horizon $T<\infty$. Assume:
\begin{enumerate}[label=(M\arabic*)]
  \item the step sizes satisfy
  \[
    s_n=\varepsilon\,\chi(n\varepsilon),
  \]
  where $\chi:[0,\infty)\to(0,\infty)$ is bounded and Lipschitz;
  \item the data stream
  \[
    Z_{n+1}=(X_{n+1},Y_{n+1})
  \]
  is i.i.d.\ with law $P$, and $|Y|\le Y_\ast$ almost surely under $P$;
  \item the feature map $\varphi$ is bounded, and $\nabla_\xi\varphi$ is bounded
  and Lipschitz in $\xi$, uniformly in $x$;
  \item the initialization law $\rho_0$ on $\Theta$ has compact support;
  \item the mixture size $K$ is fixed and finite;
  \item every component uses the softplus scale link
  $\psi_{\rm sp}(s)=\lambda_{\min}+\log(1+e^s)$ with $\lambda_{\min}>0$.
\end{enumerate}
\end{assumption}

\begin{remark}[Role of the bounded-support assumptions]
The bounded response assumption $|Y|\le Y_\ast$ and the global boundedness
assumptions on $\varphi$ and $\nabla_\xi\varphi$ are imposed here only to place
the MDN dynamics inside the bounded-support finite-horizon SGD mean-field
framework of Mei--Misiakiewicz--Montanari~\cite{MeiMisiakiewiczMontanari2019}.
They are not intended as intrinsic assumptions on Gaussian MDNs. Relaxing them
would require moment or coercivity estimates that are outside the scope of this
note.
\end{remark}

Write
\[
  K_\varphi:=\sup_{\xi,x} |\varphi(\xi,x)|,
  \qquad
  K_\nabla:=\sup_{\xi,x}\norm{\nabla_\xi\varphi(\xi,x)},
\]
and let $L_\varphi$ and $L_\nabla$ be Lipschitz constants in $\xi$ for
$\varphi$ and $\nabla_\xi\varphi$.

\begin{proposition}[Raw drift formula for the Gaussian MDN]
\label{prop:mdn-raw-drift}
Assume Assumption~\ref{ass:mdn-sgd}. For a measure $\rho$ define
\begin{align*}
  u_{k,\rho}(x,y)
  &:=
  \pi_{k,\rho}(x)-r_{k,\rho}(x,y),
  \\
  v_{1,k,\rho}(x,y)
  &:=
  r_{k,\rho}(x,y)\,q_1(\eta_{k,\rho}(x);y),
  \\
  v_{2,k,\rho}(x,y)
  &:=
  r_{k,\rho}(x,y)\,q_2(\eta_{k,\rho}(x);y).
\end{align*}
Then the mean-field-scaled population gradient flow associated with the MDN risk
has drift $b=(b_\beta,b_a,b_c,b_\xi)$ given componentwise by
\begin{align}
  b_{\beta_k}(\theta,\rho)
  &:=
  -
  \E\big[u_{k,\rho}(X,Y)\,\varphi(\xi,X)\big],
  \label{eq:mdn-drift-beta}
  \\
  b_{a_k}(\theta,\rho)
  &:=
  -
  \E\big[v_{1,k,\rho}(X,Y)\,\varphi(\xi,X)\big],
  \label{eq:mdn-drift-a}
  \\
  b_{c_k}(\theta,\rho)
  &:=
  \E\big[\psi_{\rm sp}'(s_{k,\rho}(X))\,v_{2,k,\rho}(X,Y)\,\varphi(\xi,X)\big],
  \label{eq:mdn-drift-c}
  \\
  b_\xi(\theta,\rho)
  &:=
  -
  \E\Bigg[
    \sum_{k=1}^K
    \Big(
      \beta_k u_{k,\rho}(X,Y)
      +
      a_k v_{1,k,\rho}(X,Y)
      -
      c_k\psi_{\rm sp}'(s_{k,\rho}(X))v_{2,k,\rho}(X,Y)
    \Big)\nabla_\xi\varphi(\xi,X)
  \Bigg].
  \label{eq:mdn-drift-xi}
\end{align}
\end{proposition}

\begin{proof}
Let
\[
  \mc{R}_N(\theta_1,\dots,\theta_N)
  :=
  \E\big[-\log p_N(Y|X)\big],
\]
where $p_N$ is defined in \eqref{eq:finite-mdn-density}. For particle $i$ and
component $k$,
\[
  \partial_{\theta_i} g_{k,N}(x)
  =
  \frac1N\bigl(e_k\varphi(\xi_i,x),0,0,\beta_{i,k}\nabla_\xi\varphi(\xi_i,x)\bigr),
\]
\[
  \partial_{\theta_i}\eta_{1,k,N}(x)
  =
  \frac1N\bigl(0,e_k\varphi(\xi_i,x),0,a_{i,k}\nabla_\xi\varphi(\xi_i,x)\bigr),
\]
and
\[
  \partial_{\theta_i}\eta_{2,k,N}(x)
  =
  -\frac1N\psi_{\rm sp}'(s_{k,N}(x))
  \bigl(0,0,e_k\varphi(\xi_i,x),c_{i,k}\nabla_\xi\varphi(\xi_i,x)\bigr),
\]
where $e_k$ is the $k$th basis vector of $\R^K$. Applying
Lemma~\ref{lem:mdn-differential-identities} and the chain rule gives
\[
  \nabla_{\theta_i}\mc{R}_N
  =
  \frac1N\bigl(G_{\beta,i},G_{a,i},G_{c,i},G_{\xi,i}\bigr),
\]
where
\begin{align*}
  G_{\beta,i}
  &:=
  \bigl(\E[u_{k,\rho_N}(X,Y)\varphi(\xi_i,X)]\bigr)_{k=1}^K,
  \\
  G_{a,i}
  &:=
  \bigl(\E[v_{1,k,\rho_N}(X,Y)\varphi(\xi_i,X)]\bigr)_{k=1}^K,
  \\
  G_{c,i}
  &:=
  \bigl(
    -\E[\psi_{\rm sp}'(s_{k,\rho_N}(X))v_{2,k,\rho_N}(X,Y)\varphi(\xi_i,X)]
  \bigr)_{k=1}^K,
  \\
  G_{\xi,i}
  &:=
  \E\!\Big[
    \sum_{k=1}^K
    \big(
      \beta_{i,k}u_{k,\rho_N}(X,Y)
      +a_{i,k}v_{1,k,\rho_N}(X,Y)
    \\
    &\hspace{8em}
      -c_{i,k}\psi_{\rm sp}'(s_{k,\rho_N}(X))v_{2,k,\rho_N}(X,Y)
    \big)\nabla_\xi\varphi(\xi_i,X)
  \Big].
\end{align*}
Multiplying by the mean-field factor $N$ in gradient flow yields the stated
drift formulas.
\end{proof}

\begin{equation}
  \partial_t \rho_t
  +
  \nabla_\theta\cdot\bigl(\rho_t\,\chi(t)b(\theta,\rho_t)\bigr)
  =
  0,
  \qquad
  \rho_{t=0}=\rho_0.
  \label{eq:mdn-sgd-pde}
\end{equation}

The corresponding one-sample stochastic drift and SGD recursion are needed in
the theorem statement, so we record them here rather than introducing them only
in the appendix.

\begin{definition}[One-sample stochastic drift for the MDN]
\label{def:mdn-sample-drift}
For $z=(x,y)\in\mc{X}\times\R$, define
\[
  \hat b(\theta,\rho;z)
  :=
  \bigl(\hat b_\beta(\theta,\rho;z),\hat b_a(\theta,\rho;z),\hat b_c(\theta,\rho;z),\hat b_\xi(\theta,\rho;z)\bigr)
\]
with componentwise coordinates
\begin{align}
  \hat b_{\beta_k}(\theta,\rho;z)
  &:=
  -
  u_{k,\rho}(x,y)\,\varphi(\xi,x),
  \label{eq:mdn-sample-drift-beta}
  \\
  \hat b_{a_k}(\theta,\rho;z)
  &:=
  -
  v_{1,k,\rho}(x,y)\,\varphi(\xi,x),
  \label{eq:mdn-sample-drift-a}
  \\
  \hat b_{c_k}(\theta,\rho;z)
  &:=
  \psi_{\rm sp}'(s_{k,\rho}(x))\,v_{2,k,\rho}(x,y)\,\varphi(\xi,x),
  \label{eq:mdn-sample-drift-c}
  \\
  \hat b_\xi(\theta,\rho;z)
  &:=
  -
  \sum_{k=1}^K
  \Big(
    \beta_k u_{k,\rho}(x,y)
    +
    a_k v_{1,k,\rho}(x,y)
    -
    c_k\psi_{\rm sp}'(s_{k,\rho}(x))v_{2,k,\rho}(x,y)
  \Big)\nabla_\xi\varphi(\xi,x).
  \label{eq:mdn-sample-drift-xi}
\end{align}
\end{definition}

\begin{proposition}[Unbiased raw stochastic update for the MDN]
\label{prop:mdn-unbiased-sample-drift}
Under Assumption~\ref{ass:mdn-sgd},
\[
  b(\theta,\rho)=\E_Z[\hat b(\theta,\rho;Z)],
\]
where $Z=(X,Y)\sim P$. Consequently the plain SGD iteration can be written
as
\begin{equation}
  \theta_i^{n+1}
  =
  \theta_i^n
  +
  s_n\,\hat b(\theta_i^n,\hat\rho_n^N;Z_{n+1}),
  \qquad
  \hat\rho_n^N:=\frac1N\sum_{i=1}^N \delta_{\theta_i^n}.
  \label{eq:mdn-sgd-update}
\end{equation}
\end{proposition}

\begin{proof}
Take expectations of
\eqref{eq:mdn-sample-drift-beta}--\eqref{eq:mdn-sample-drift-xi} with respect
to $Z=(X,Y)\sim P$ and compare with
\eqref{eq:mdn-drift-beta}--\eqref{eq:mdn-drift-xi}.
\end{proof}

The rest of the training-side work is technical verification rather than a new
conceptual step. Appendix~\ref{app:training-verification} records the
MDN-specific confinement, drift
regularity, and one-step SGD bounds needed to place the present drift inside the
bounded-support finite-horizon SGD mean-field framework of
Mei--Misiakiewicz--Montanari~\cite{MeiMisiakiewiczMontanari2019} and the
shared-path route used in LMC work such as Ferbach et al.~\cite{FerbachEtAl2024}.

\begin{theorem}[Training-side theorem: MDN instantiation of the bounded-support SGD mean-field framework]
\label{thm:mdn-sgd-mean-field}
Assume Assumption~\ref{ass:mdn-sgd}. Consider the plain SGD iterates
\eqref{eq:mdn-sgd-update} and let $\rho_t$ be the deterministic measure-valued
path solving \eqref{eq:mdn-sgd-pde}. Then the bounded-support finite-horizon SGD
mean-field theory of Mei--Misiakiewicz--Montanari, instantiated with the MDN
drift above, yields:
\begin{enumerate}[label=(\roman*)]
  \item the PDE \eqref{eq:mdn-sgd-pde} is well posed on $[0,T]$;
  \item there exists an error term
  $\delta_{N,\varepsilon}^{\rm mdn}(T)\to 0$ as $N\to\infty$ and
  $\varepsilon\downarrow 0$ such that
  \[
    \sup_{n\le T/\varepsilon}
    W_1\!\bigl(\hat\rho_n^N,\rho_{n\varepsilon}\bigr)
    \le
    \delta_{N,\varepsilon}^{\rm mdn}(T)
  \]
  with high probability;
  \item in bounded-support regimes, one obtains the same qualitative structure
  as in the Mei/Ferbach four-dynamics argument: a Monte-Carlo term of order
  $N^{-1/2}$ together with discretization and stochastic-gradient terms of
  order $\varepsilon$ and $\sqrt{\varepsilon}$.
\end{enumerate}
\end{theorem}

\begin{proof}[Framework-instantiation proof sketch]
Proposition~\ref{prop:mdn-unbiased-sample-drift} identifies the SGD recursion,
and Appendix~\ref{app:training-verification} proves
Propositions~\ref{prop:mdn-confinement},
\ref{prop:mdn-drift-regularity},
\ref{prop:mdn-sgd-discrete-confinement}, and
\ref{prop:mdn-sample-drift-regularity}. These are exactly the MDN-specific
bounded-support hypotheses needed to invoke the finite-horizon SGD mean-field
machinery of Mei--Misiakiewicz--Montanari and its shared-path use in the LMC
framework of Ferbach et al. In particular,
Proposition~\ref{prop:mdn-drift-regularity} gives the well-posedness of the
characteristic flow on the confinement set, and the remaining propositions
provide the boundedness, unbiasedness, and one-step regularity estimates
required by the general theorem. The bounded Lipschitz schedule $\chi$ enters
only as a time-inhomogeneous scalar prefactor in the drift and is handled by the
same finite-horizon argument after the standard time-change reduction. The
result is therefore a framework instantiation rather than a standalone new
convergence proof.
\end{proof}

\begin{corollary}[Training-side corollary: shared-law common path for SGD in Gaussian MDNs]
\label{cor:mdn-sgd-common-path}
Assume Assumption~\ref{ass:mdn-sgd}. Let $\hat\rho_{A,n}^N,\hat\rho_{B,n}^N$ be
two independent width-$N$ SGD runs of \eqref{eq:mdn-sgd-update}, both
initialized i.i.d.\ from the same law $\rho_0$ and trained over the same
horizon $[0,T]$. Then
\[
  \sup_{n\le T/\varepsilon}
  W_1\!\bigl(\hat\rho_{A,n}^N,\hat\rho_{B,n}^N\bigr)
  \le
  2\,\delta_{N,\varepsilon}^{\rm mdn}(T)
\]
with high probability. In particular,
\[
  \sup_{n\le T/\varepsilon}
  W_1\!\bigl(\hat\rho_{A,n}^N,\hat\rho_{B,n}^N\bigr)\to 0
\]
in probability as $N\to\infty$ and $\varepsilon\downarrow 0$.
\end{corollary}

\begin{proof}
Apply Theorem~\ref{thm:mdn-sgd-mean-field} to each run and use the triangle
inequality.
\end{proof}

\begin{remark}[How to read the SGD common-path theorem]
Theorem~\ref{thm:mdn-sgd-mean-field} and
Corollary~\ref{cor:mdn-sgd-common-path} should be read as \emph{repeatability in
a chosen labeled gauge}. For a fixed initialization law $\rho_0$ and fixed
finite horizon $[0,T]$, the wide/small-step limit of plain SGD is
deterministic in labeled parameter space. What this does \emph{not} provide is
a canonical component decomposition of the underlying predictor.
\end{remark}

\section{Why the \texorpdfstring{$K=1$}{K=1} Synthesis Route Breaks for \texorpdfstring{$K>1$}{K>1}}

\begin{definition}[Labeled gauge, predictive equivalence, and global permutation]
For a $K$-component MDN written in the raw coordinates
$(\beta,a,c,\xi)$, the \emph{labeled gauge} means that the component labels
$k=1,\dots,K$ are kept fixed and compared coordinatewise. Two $K$-component
MDN predictors are \emph{predictively equivalent} if they induce the same
conditional density $p(y|x)$ for every $(x,y)$. The \emph{mixture gauge}
refers to the resulting non-identifiability inside such a predictive
equivalence class: multiple labeled component decompositions can represent the
same predictor. A \emph{global permutation} acts on a predictor by applying the
same permutation of $\{1,\dots,K\}$ to the component indices for every $x$.
\end{definition}

The common-path theorem fixes the dynamics only in the labeled gauge. It does
not choose a canonical representative inside the mixture gauge class of the
corresponding predictor. For $K=1$ this distinction is absent or much smaller;
for $K>1$ it is the new synthesis obstruction.

\noindent
For the obstruction example we now specialize to scalar inputs, $d=1$.

\begin{theorem}[Global-permutation linear synthesis can fail already at the head level]
\label{thm:mdn-obstruction}
Fix $\sigma^2>0$ and let $\phi_\sigma(\cdot-m)$ denote the Gaussian density
with mean $m$ and variance $\sigma^2$. Let $I\subseteq \R$ be any input domain
containing both negative and positive points, and consider the two $K=2$ MDN
predictors on $I$
\[
  p_A(y|x)
  :=
  \frac12 \phi_\sigma(y-x)+\frac12 \phi_\sigma(y+x),
\]
and
\[
  p_B(y|x)
  :=
  \frac12 \phi_\sigma(y-|x|)+\frac12 \phi_\sigma(y+|x|).
\]
Then:
\begin{enumerate}[label=(\roman*)]
  \item $p_A(y|x)=p_B(y|x)$ for every $(x,y)$, so the two predictors are
  predictively equivalent;
  \item for any global permutation $\tau\in S_2$, the arithmetic midpoint in
  the labeled natural-coordinate output-head fields of $A$ and $\tau\cdot B$
  produces a predictor $\bar p_\tau$ such that either
  \[
    \bar p_\tau(\cdot|x)\neq p_A(\cdot|x)\qquad \forall x<0,
  \]
  or
  \[
    \bar p_\tau(\cdot|x)\neq p_A(\cdot|x)\qquad \forall x>0;
  \]
  \item if the data law gives positive mass to both $\{x<0\}$ and $\{x>0\}$,
  and $Y|X\sim p_A(\cdot|X)$, then the population negative log-likelihood
  satisfies
  \[
    \E[-\log \bar p_\tau(Y|X)]
    >
    \E[-\log p_A(Y|X)]
  \]
  for every global permutation $\tau$.
\end{enumerate}
\end{theorem}

\begin{proof}[Proof sketch]
For $x\ge 0$ the two representations agree componentwise, while for $x<0$ they
agree only after swapping the two components. Therefore any fixed global
labeling disagrees with one half of the input domain. On that half, the
component midpoint collapses to the single Gaussian $\phi_\sigma(y)$, which
differs from the common target mixture and yields a strictly positive KL gap.
Appendix~\ref{app:obstruction-proof} records the full calculation.
\end{proof}

\begin{corollary}[Direct extension of the $K=1$ synthesis premise fails for plain MDNs]
\label{cor:k1-breaks-at-k2}
Any LMC strategy whose synthesis step requires predictively equivalent endpoint
representations to become globally alignable after a single component
permutation can fail already for plain $K=2$ MDNs. Consequently, the direct
$K=1\to K>1$ extension breaks at the synthesis premise, not at the training-side
common-path premise.
\end{corollary}

\begin{proof}
For $K=1$ there is no nontrivial mixture decomposition to identify. For $K=2$,
Theorem~\ref{thm:mdn-obstruction} provides a pair of predictively equivalent
heads for which every global-permutation midpoint has strictly larger
population loss. Hence the naive $K=1$ synthesis step does not survive in plain
mixture models.
\end{proof}

\begin{remark}[Interpretation for future LMC work]
Theorem~\ref{thm:mdn-obstruction} does \emph{not} contradict
Corollary~\ref{cor:mdn-sgd-common-path}. The finite-horizon common-path theorem
says that, for a fixed initialization law and fixed labeled parameter space, the
wide/small-step limit dynamics are deterministic. The obstruction theorem says
that the labeled parameter space itself is not canonically quotiented by global
permutation. Thus the natural expectation for plain MDNs is:
\begin{quote}
In the joint wide/small-step mean-field limit and at fixed finite horizon,
run-to-run gauge randomness should be suppressed for a fixed initialization law;
at finite width and with SGD/noise, different microscopic asymmetries can
select different representatives, and the plain global-permutation route to
linear mode connectivity can fail.
\end{quote}
\end{remark}

\section{Finite-Width Interpretation}

The gauge is not created by finite width: the mixture non-identifiability
already exists at the level of the predictor class. Finite width and SGD noise
matter because different runs can select different representatives of the same
predictive class. Accordingly, the finite-width discrepancy should be read as
having a transverse component and a gauge component.

\begin{definition}[Finite-width synthesis barrier]
\label{def:barrier}
Let $\Theta_A^N,\Theta_B^N$ be two trained width-$N$ particle systems after a
chosen neuron alignment procedure. Their linear-interpolation barrier is
\[
  \mathrm{Bar}_N(\Theta_A^N,\Theta_B^N)
  :=
  \sup_{\alpha\in[0,1]}
  \mc{R}_N\big((1-\alpha)\Theta_A^N+\alpha \Theta_B^N\big)
  -
  \max\{\mc{R}_N(\Theta_A^N),\mc{R}_N(\Theta_B^N)\},
\]
where $\mc{R}_N$ is the width-$N$ population MDN risk on particle systems.
\end{definition}

\begin{conjecture}[Informal gauge-dominance picture]
After separating the ordinary finite-width tracking error from the
run-to-run discrepancy, the remaining synthesis barrier for plain MDNs is
dominated by the component of that discrepancy lying along mixture-gauge
directions. For $K=1$ the corresponding nontrivial mixture gauge is absent, so
only the ordinary finite-width error remains.
\end{conjecture}

\begin{remark}[Formal template]
Appendix~\ref{app:gauge-template} records one formal stable/gauge decomposition
template for turning the previous conjectural picture into a theorem. It is
kept in the appendix because it is a proof-building device rather than part of
the main narrative of the note.
\end{remark}

\section{Implications for the Research Program}

The present note supports the following research direction.

\begin{enumerate}[label=(\roman*)]
  \item The \emph{training-side} premise used in the $K=1$ LMC argument still
  survives for plain MDNs. In the current note it is verified in the same
  finite-horizon plain-SGD scaling, but only as a deterministic statement
  in a chosen labeled gauge.
  \item The direct extension from $K=1$ to $K>1$ fails at the
  \emph{synthesis-side} premise, not at the mean-field premise.
  \item Consequently, an eventual $K>1$ analogue of the $K=1$ theorem must
  either prove that a richer quotient of MDN parameter space is compatible with
  synthesis, or avoid the issue by changing the model class.
\end{enumerate}

The clean narrative is therefore:
\begin{quote}
For $K=1$, one can hope to run the full chain from finite-horizon common path to
aligned interpolation to LMC. For $K>1$, the wide-limit common path still looks
plausible, but the chain can break before interpolation because the relevant
quotient of MDN parameter space is no longer captured by global permutation.
\end{quote}

\appendix

\section{Training-Side Verification}
\label{app:training-verification}

This appendix records the MDN-specific confinement, regularity, and one-step SGD
estimates used in the proof sketch of
Theorem~\ref{thm:mdn-sgd-mean-field}. None of these are conceptually the main
point of the note; they are the bookkeeping needed to fit the present MDN drift
into the bounded-support finite-horizon SGD framework of
Mei--Misiakiewicz--Montanari and Ferbach et al.

\begin{proposition}[Time-inhomogeneous finite-horizon confinement for the MDN flow]
\label{prop:mdn-confinement}
Assume Assumption~\ref{ass:mdn-sgd} and let
\[
  \chi_{\max}:=\sup_{t\in[0,T]}\chi(t)<\infty.
\]
Define the initial bounds
\begin{align*}
  B_0
  &:=
  \sup_{\theta\in\supp(\rho_0)}\max_{1\le k\le K}|\beta_k|,
  \\
  A_0
  &:=
  \sup_{\theta\in\supp(\rho_0)}\max_{1\le k\le K}|a_k|,
  \\
  C_0
  &:=
  \sup_{\theta\in\supp(\rho_0)}\max_{1\le k\le K}|c_k|,
  \\
  R_0
  &:=
  \sup_{\theta\in\supp(\rho_0)}\norm{\xi}.
\end{align*}
Let $B_t^{\rm sgd},A_t^{\rm sgd},C_t^{\rm sgd},R_t^{\rm sgd}$ solve
\begin{align}
  \dot B_t^{\rm sgd}
  &=
  \chi_{\max}K_\varphi,
  \qquad
  B_{t=0}^{\rm sgd}=B_0,
  \label{eq:B-envelope-mdn-sgd}
  \\
  \Gamma_{1,t}^{\rm sgd}
  &:=
  \frac{K_\varphi A_t^{\rm sgd}}{2\lambda_{\min}}+Y_\ast,
  \label{eq:Gamma1-mdn-sgd}
  \\
  \dot A_t^{\rm sgd}
  &=
  \chi_{\max}K_\varphi \Gamma_{1,t}^{\rm sgd},
  \qquad
  A_{t=0}^{\rm sgd}=A_0,
  \label{eq:A-envelope-mdn-sgd}
  \\
  \Gamma_{2,t}^{\rm sgd}
  &:=
  \frac{K_\varphi^2 (A_t^{\rm sgd})^2}{4\lambda_{\min}^2}
  +
  \frac{1}{2\lambda_{\min}}
  +
  Y_\ast^2,
  \label{eq:Gamma2-mdn-sgd}
  \\
  \dot C_t^{\rm sgd}
  &=
  \chi_{\max}K_\varphi \Gamma_{2,t}^{\rm sgd},
  \qquad
  C_{t=0}^{\rm sgd}=C_0,
  \label{eq:C-envelope-mdn-sgd}
  \\
  \dot R_t^{\rm sgd}
  &=
  \chi_{\max}K K_\nabla
  \big(
    B_t^{\rm sgd}
    +
    A_t^{\rm sgd}\Gamma_{1,t}^{\rm sgd}
    +
    C_t^{\rm sgd}\Gamma_{2,t}^{\rm sgd}
  \big),
  \qquad
  R_{t=0}^{\rm sgd}=R_0.
  \label{eq:R-envelope-mdn-sgd}
\end{align}
Then these envelopes remain finite on $[0,T]$. Every characteristic of the
time-inhomogeneous mean-field flow
\[
  \dot\theta(t)=\chi(t)b(\theta(t),\rho_t)
\]
started from $\supp(\rho_0)$ satisfies
\[
  \begin{aligned}
    \max_k|\beta_k(t)|&\le B_t^{\rm sgd},
    \qquad
    \max_k|a_k(t)|\le A_t^{\rm sgd},
    \\
    \max_k|c_k(t)|&\le C_t^{\rm sgd},
    \qquad
    \norm{\xi(t)}\le R_t^{\rm sgd},
    \qquad t\in[0,T].
  \end{aligned}
\]
Consequently, with
\[
  \begin{aligned}
    B_t^{\Theta,\rm sgd}
    &:=
    \left\{
      (\beta,a,c,\xi):
      \max_k|\beta_k|\le B_t^{\rm sgd},\;
      \max_k|a_k|\le A_t^{\rm sgd},
    \right.
    \\
    &\qquad\left.
      \max_k|c_k|\le C_t^{\rm sgd},\;
      \norm{\xi}\le R_t^{\rm sgd}
    \right\},
    \qquad t\in[0,T].
  \end{aligned}
\]
every solution remains inside
\[
  B_t^{\Theta,\rm sgd}\subseteq B_T^{\Theta,\rm sgd},
\]
and the corresponding fields satisfy
\[
  |g_{k,\rho_t}(x)|\le K_\varphi B_t^{\rm sgd},
  \qquad
  |\eta_{1,k,\rho_t}(x)|\le K_\varphi A_t^{\rm sgd},
  \qquad
  |s_{k,\rho_t}(x)|\le K_\varphi C_t^{\rm sgd}
\]
for all $k,x,t\in[0,T]$.
\end{proposition}

\begin{proof}
Because $0\le \pi_{k,\rho}\le 1$ and $0\le r_{k,\rho}\le 1$,
\[
  |u_{k,\rho}(x,y)|\le 1.
\]
Also $\eta_{2,k,\rho}(x)\le -\lambda_{\min}$ and
$|\eta_{1,k,\rho}(x)|\le K_\varphi A_t^{\rm sgd}$, hence
\[
  |q_1(\eta_{k,\rho}(x);y)|\le \Gamma_{1,t}^{\rm sgd},
  \qquad
  |q_2(\eta_{k,\rho}(x);y)|\le \Gamma_{2,t}^{\rm sgd}.
\]
Since $0<\psi_{\rm sp}'<1$, the drift formulas
\eqref{eq:mdn-drift-beta}--\eqref{eq:mdn-drift-xi} imply
\[
  |\dot\beta_k(t)|\le \chi_{\max}K_\varphi,
  \qquad
  |\dot a_k(t)|\le \chi_{\max}K_\varphi\Gamma_{1,t}^{\rm sgd},
  \qquad
  |\dot c_k(t)|\le \chi_{\max}K_\varphi\Gamma_{2,t}^{\rm sgd},
\]
and
\[
  \norm{\dot\xi(t)}
  \le
  \chi_{\max}K K_\nabla
  \big(
    B_t^{\rm sgd}
    +
    A_t^{\rm sgd}\Gamma_{1,t}^{\rm sgd}
    +
    C_t^{\rm sgd}\Gamma_{2,t}^{\rm sgd}
  \big).
\]
Comparison for scalar ODEs yields the claimed bounds. The field bounds follow
from the estimates
\[
  |g_{k,\rho_t}(x)|
  \le
  \int |\beta_k|\,|\varphi(\xi,x)|\,\rho_t(d\theta)
  \le
  K_\varphi B_t^{\rm sgd},
\]
and similarly for $\eta_{1,k,\rho_t}$ and $s_{k,\rho_t}$.
\end{proof}

\begin{proposition}[Drift regularity on the finite-horizon confinement set]
\label{prop:mdn-drift-regularity}
Assume Assumption~\ref{ass:mdn-sgd}. There exists a constant
$L_T^{\rm mdn}<\infty$ such that for all
$\theta,\bar\theta\in B_T^{\Theta,\rm sgd}$ and all probability measures
$\rho,\nu$ supported on $B_T^{\Theta,\rm sgd}$,
\begin{equation}
  \norm{b(\theta,\rho)-b(\bar\theta,\nu)}
  \le
  L_T^{\rm mdn}\big(\norm{\theta-\bar\theta}+W_1(\rho,\nu)\big),
  \label{eq:mdn-drift-lipschitz}
\end{equation}
and
\[
  \sup_{\theta\in B_T^{\Theta,\rm sgd},\ \supp(\rho)\subseteq B_T^{\Theta,\rm sgd}}
  \norm{b(\theta,\rho)}
  \le
  L_T^{\rm mdn}.
\]
\end{proposition}

\begin{proof}
On $B_T^{\Theta,\rm sgd}$, the maps
\[
  \theta\mapsto \beta_k\varphi(\xi,x),
  \qquad
  \theta\mapsto a_k\varphi(\xi,x),
  \qquad
  \theta\mapsto c_k\varphi(\xi,x)
\]
are uniformly bounded and Lipschitz in $\theta$, uniformly in $x$ and $k$.
Therefore, for every $x$ and $k$,
\[
  |g_{k,\rho}(x)-g_{k,\nu}(x)|
  +
  |\eta_{1,k,\rho}(x)-\eta_{1,k,\nu}(x)|
  +
  |s_{k,\rho}(x)-s_{k,\nu}(x)|
  \le
  C_T^{(1)} W_1(\rho,\nu)
\]
for some finite constant $C_T^{(1)}$.

By Proposition~\ref{prop:mdn-confinement},
\[
  |g_{k,\rho}(x)|\le G_T^{\rm sgd}:=K_\varphi B_T^{\rm sgd},
  \qquad
  |\eta_{1,k,\rho}(x)|\le M_T^{\rm sgd}:=K_\varphi A_T^{\rm sgd},
  \qquad
  -\Lambda_T^{\rm sgd}\le \eta_{2,k,\rho}(x)\le -\lambda_{\min},
\]
where $\Lambda_T^{\rm sgd}:=\psi_{\rm sp}(K_\varphi C_T^{\rm sgd})$. Hence the
component score
\[
  z_{k,\rho}(x,y)
  :=
  g_{k,\rho}(x)
  +
  \eta_{1,k,\rho}(x)y
  +
  \eta_{2,k,\rho}(x)y^2
  -
  A_{\rm G}(\eta_{k,\rho}(x))
\]
is uniformly bounded and Lipschitz in $(g,\eta)$ on the compact set
\[
  [-G_T^{\rm sgd},G_T^{\rm sgd}]^K
  \times
  \Big(
    [-M_T^{\rm sgd},M_T^{\rm sgd}]
    \times
    [-\Lambda_T^{\rm sgd},-\lambda_{\min}]
  \Big)^K
  \times
  [-Y_\ast,Y_\ast].
\]
Since the softmax map on $\R^K$ is globally Lipschitz, both the gate map
$g\mapsto \pi(g)$ and the responsibility map $z\mapsto \softmax(z)$ are
Lipschitz on this compact set. Therefore
\[
  (\rho,x,y)\mapsto u_{k,\rho}(x,y),
  \qquad
  (\rho,x,y)\mapsto v_{1,k,\rho}(x,y),
  \qquad
  (\rho,x,y)\mapsto v_{2,k,\rho}(x,y)
\]
are uniformly bounded and Lipschitz in $\rho$ on the confinement set. Inserting
these bounds into \eqref{eq:mdn-drift-beta}--\eqref{eq:mdn-drift-xi}, and using
the boundedness and Lipschitz continuity of $\varphi$ and $\nabla_\xi\varphi$,
gives \eqref{eq:mdn-drift-lipschitz}. The uniform bound on $b$ follows from the
same estimates with $\bar\theta=\theta$ and $\rho=\nu$.
\end{proof}

\begin{proposition}[Discrete-time finite-horizon confinement for MDN SGD]
\label{prop:mdn-sgd-discrete-confinement}
Assume Assumption~\ref{ass:mdn-sgd}. Define deterministic sequences
$B_n^\varepsilon,A_n^\varepsilon,C_n^\varepsilon,R_n^\varepsilon$ by
\begin{align}
  B_{n+1}^\varepsilon
  &=
  B_n^\varepsilon + s_n K_\varphi,
  \qquad
  B_0^\varepsilon=B_0,
  \label{eq:mdn-discrete-B-envelope}
  \\
  A_{n+1}^\varepsilon
  &=
  A_n^\varepsilon
  +
  s_n K_\varphi
  \left(
    \frac{K_\varphi A_n^\varepsilon}{2\lambda_{\min}}+Y_\ast
  \right),
  \qquad
  A_0^\varepsilon=A_0,
  \label{eq:mdn-discrete-A-envelope}
  \\
  C_{n+1}^\varepsilon
  &=
  C_n^\varepsilon
  +
  s_n K_\varphi
  \left(
    \frac{K_\varphi^2 (A_n^\varepsilon)^2}{4\lambda_{\min}^2}
    +
    \frac{1}{2\lambda_{\min}}
    +
    Y_\ast^2
  \right),
  \qquad
  C_0^\varepsilon=C_0,
  \label{eq:mdn-discrete-C-envelope}
  \\
  R_{n+1}^\varepsilon
  &=
  R_n^\varepsilon
  +
  s_n K K_\nabla
  \Biggl[
    B_n^\varepsilon
    +
    A_n^\varepsilon
    \left(
      \frac{K_\varphi A_n^\varepsilon}{2\lambda_{\min}}+Y_\ast
    \right)
    \nonumber\\
  &\qquad\qquad\qquad
    +
    C_n^\varepsilon
    \left(
      \frac{K_\varphi^2 (A_n^\varepsilon)^2}{4\lambda_{\min}^2}
      +
      \frac{1}{2\lambda_{\min}}
      +
      Y_\ast^2
    \right)
  \Biggr],
  \qquad
  R_0^\varepsilon=R_0.
  \label{eq:mdn-discrete-R-envelope}
\end{align}
Then every SGD iterate of \eqref{eq:mdn-sgd-update} satisfies
\[
  \begin{aligned}
    \max_{1\le k\le K}|\beta_{i,k}^n|&\le B_n^\varepsilon,
    \qquad
    \max_{1\le k\le K}|a_{i,k}^n|\le A_n^\varepsilon,
    \\
    \max_{1\le k\le K}|c_{i,k}^n|&\le C_n^\varepsilon,
    \qquad
    \norm{\xi_i^n}\le R_n^\varepsilon,
    \qquad
    \forall i,\ \forall n\le T/\varepsilon.
  \end{aligned}
\]
Moreover, the discrete envelopes are uniformly bounded by the continuous
majorants from Proposition~\ref{prop:mdn-confinement}; in particular every SGD
iterate remains in $B_T^{\Theta,\rm sgd}$.
\end{proposition}

\begin{proof}
We argue by induction on $n$. Assume
\[
  \max_{1\le k\le K}|\beta_{i,k}^n|\le B_n^\varepsilon,
  \qquad
  \max_{1\le k\le K}|a_{i,k}^n|\le A_n^\varepsilon,
  \qquad
  \max_{1\le k\le K}|c_{i,k}^n|\le C_n^\varepsilon,
  \qquad
  \norm{\xi_i^n}\le R_n^\varepsilon
\]
for all $i$. Then, for every component index $k$,
\[
  |g_{k,\hat\rho_n^N}(x)|
  \le
  \frac1N\sum_{i=1}^N |\beta_{i,k}^n|\,|\varphi(\xi_i^n,x)|
  \le
  K_\varphi B_n^\varepsilon,
\]
\[
  |\eta_{1,k,\hat\rho_n^N}(x)|
  \le
  \frac1N\sum_{i=1}^N |a_{i,k}^n|\,|\varphi(\xi_i^n,x)|
  \le
  K_\varphi A_n^\varepsilon,
\]
and
\[
  |s_{k,\hat\rho_n^N}(x)|
  \le
  \frac1N\sum_{i=1}^N |c_{i,k}^n|\,|\varphi(\xi_i^n,x)|
  \le
  K_\varphi C_n^\varepsilon.
\]
Proposition~\ref{prop:mdn-domain} gives
$\eta_{2,k,\hat\rho_n^N}(x)\le -\lambda_{\min}$ for all $k$. Hence
\[
  |u_{k,\hat\rho_n^N}(x,y)|\le 1,
  \qquad
  |v_{1,k,\hat\rho_n^N}(x,y)|
  \le
  \frac{K_\varphi A_n^\varepsilon}{2\lambda_{\min}}+Y_\ast,
\]
and
\[
  |v_{2,k,\hat\rho_n^N}(x,y)|
  \le
  \frac{K_\varphi^2 (A_n^\varepsilon)^2}{4\lambda_{\min}^2}
  +
  \frac{1}{2\lambda_{\min}}
  +
  Y_\ast^2.
\]
Inserting these bounds into the one-sample update
\eqref{eq:mdn-sgd-update} yields
\[
  \max_{1\le k\le K}|\beta_{i,k}^{n+1}|
  \le
  B_{n+1}^\varepsilon,
  \qquad
  \max_{1\le k\le K}|a_{i,k}^{n+1}|
  \le
  A_{n+1}^\varepsilon,
  \qquad
  \max_{1\le k\le K}|c_{i,k}^{n+1}|
  \le
  C_{n+1}^\varepsilon,
\]
and
\[
  \norm{\xi_i^{n+1}}
  \le
  R_{n+1}^\varepsilon.
\]
This closes the induction.

Since $s_n\le \varepsilon\chi_{\max}$ and the right-hand sides of
\eqref{eq:mdn-discrete-B-envelope}--\eqref{eq:mdn-discrete-R-envelope} are
nondecreasing in the envelopes, the standard Euler-vs-ODE comparison gives
\[
  B_n^\varepsilon\le B_{n\varepsilon}^{\rm sgd},
  \qquad
  A_n^\varepsilon\le A_{n\varepsilon}^{\rm sgd},
  \qquad
  C_n^\varepsilon\le C_{n\varepsilon}^{\rm sgd},
  \qquad
  R_n^\varepsilon\le R_{n\varepsilon}^{\rm sgd}
\]
for all $n\le T/\varepsilon$.
\end{proof}

\begin{proposition}[Uniform stochastic-drift regularity for the MDN]
\label{prop:mdn-sample-drift-regularity}
Assume Assumption~\ref{ass:mdn-sgd}. There exists
$\widehat L_T^{\rm mdn}<\infty$ such that for every sample $z=(x,y)$ in the
support of $P$, all $\theta,\bar\theta\in B_T^{\Theta,\rm sgd}$, and all
probability measures $\rho,\nu$ supported on $B_T^{\Theta,\rm sgd}$,
\begin{align}
  \norm{\hat b(\theta,\rho;z)}
  &\le
  \widehat L_T^{\rm mdn},
  \label{eq:mdn-sample-drift-bound}
  \\
  \norm{\hat b(\theta,\rho;z)-\hat b(\bar\theta,\nu;z)}
  &\le
  \widehat L_T^{\rm mdn}
  \bigl(\norm{\theta-\bar\theta}+W_1(\rho,\nu)\bigr).
  \label{eq:mdn-sample-drift-lipschitz}
\end{align}
\end{proposition}

\begin{proof}
On $B_T^{\Theta,\rm sgd}$, the maps
\[
  \theta\mapsto \beta_k\varphi(\xi,x),
  \qquad
  \theta\mapsto a_k\varphi(\xi,x),
  \qquad
  \theta\mapsto c_k\varphi(\xi,x)
\]
are uniformly bounded and Lipschitz in $\theta$, uniformly in $x$ and $k$.
Hence
\[
  |g_{k,\rho}(x)-g_{k,\nu}(x)|
  +
  |\eta_{1,k,\rho}(x)-\eta_{1,k,\nu}(x)|
  +
  |s_{k,\rho}(x)-s_{k,\nu}(x)|
  \le
  C_T^{(2)}W_1(\rho,\nu)
\]
for some finite constant $C_T^{(2)}$. By Proposition~\ref{prop:mdn-confinement},
all relevant fields lie in the compact set
\[
  G_T^{\rm sgd}:=K_\varphi B_T^{\rm sgd},
  \qquad
  M_T^{\rm sgd}:=K_\varphi A_T^{\rm sgd},
  \qquad
  \Lambda_T^{\rm sgd}:=\psi_{\rm sp}(K_\varphi C_T^{\rm sgd}),
\]
\[
  [-G_T^{\rm sgd},G_T^{\rm sgd}]^K
  \times
  \Big(
    [-M_T^{\rm sgd},M_T^{\rm sgd}]
    \times
    [-\Lambda_T^{\rm sgd},-\lambda_{\min}]
  \Big)^K
  \times
  [-Y_\ast,Y_\ast].
\]
On this set the component score map is bounded and Lipschitz, and the softmax
map is globally Lipschitz. Therefore
\[
  (\rho,x,y)\mapsto u_{k,\rho}(x,y),
  \qquad
  (\rho,x,y)\mapsto v_{1,k,\rho}(x,y),
  \qquad
  (\rho,x,y)\mapsto v_{2,k,\rho}(x,y)
\]
are uniformly bounded and Lipschitz in $\rho$. Inserting these bounds into
\eqref{eq:mdn-sample-drift-beta}--\eqref{eq:mdn-sample-drift-xi}, and using the
boundedness and Lipschitz continuity of $\varphi$ and $\nabla_\xi\varphi$,
gives \eqref{eq:mdn-sample-drift-bound} and
\eqref{eq:mdn-sample-drift-lipschitz}.
\end{proof}

\section{Proof of Theorem~\ref{thm:mdn-obstruction}}
\label{app:obstruction-proof}

Item (i) is immediate: for $x\ge 0$, the two formulas coincide term by term,
while for $x<0$ they coincide after swapping the two components.

For item (ii), write the two labeled representations in natural coordinates.
Both predictors have equal gate weights $\pi_1=\pi_2=1/2$ and constant second
natural coordinates
\[
  \eta_{2,1}=\eta_{2,2}=-\frac{1}{2\sigma^2}.
\]
The first natural coordinates are
\[
  \eta_{1,1}^A(x)=\frac{x}{\sigma^2},
  \qquad
  \eta_{1,2}^A(x)=-\frac{x}{\sigma^2},
\]
and
\[
  \eta_{1,1}^B(x)=\frac{|x|}{\sigma^2},
  \qquad
  \eta_{1,2}^B(x)=-\frac{|x|}{\sigma^2}.
\]
If $\tau$ is the identity, then for every $x<0$ the midpoint satisfies
\[
  \bar\eta_{1,1}(x)
  =
  \frac{1}{2}\left(\frac{x}{\sigma^2}+\frac{-x}{\sigma^2}\right)=0,
  \qquad
  \bar\eta_{1,2}(x)
  =
  \frac{1}{2}\left(\frac{-x}{\sigma^2}+\frac{x}{\sigma^2}\right)=0.
\]
Hence $\bar p_{\mathrm{id}}(y|x)=\phi_\sigma(y)$ for every $x<0$, which differs
from $p_A(y|x)$ whenever $x\ne 0$. If $\tau$ is the transposition, the same
collapse occurs for every $x>0$. Thus every global permutation fails on one of
the two sign regions: either on all $x<0$ or on all $x>0$.

For item (iii), define the midpoint excess risk
\[
  \Delta_\tau
  :=
  \E\big[-\log \bar p_\tau(Y|X)\big]
  -
  \E\big[-\log p_A(Y|X)\big].
\]
Since $Y|X\sim p_A(\cdot|X)$,
\[
  \Delta_\tau
  =
  \E\Big[
    \mathrm{KL}\big(p_A(\cdot|X)\,\|\,\bar p_\tau(\cdot|X)\big)
  \Big].
\]
By item (ii), the conditional densities differ on a set of $x$ values with
positive $P_X$-measure, so the conditional KL divergence is strictly positive
there. Therefore $\Delta_\tau>0$.

\section{Formal Gauge-Dominance Template}
\label{app:gauge-template}

This appendix records one formal theorem template for the strongest possible
finite-width interpretation of the main obstruction story. It is not needed for
the main narrative; it is included as a proof-building scaffold for future
work.

\begin{assumption}[Stable/gauge decomposition at the reference path]
\label{ass:gauge-dominance}
Fix a terminal time $T<\infty$ and let $\rho_T$ be the deterministic mean-field
law from Theorem~\ref{thm:mdn-sgd-mean-field}. Assume:
\begin{enumerate}[label=(GD\arabic*)]
  \item there is a local decomposition of perturbations around $\rho_T$ into
  stable directions and gauge directions,
  \[
    \Delta = \Delta_{\mathrm{st}}+\Delta_{\mathrm{g}},
    \qquad
    \Delta_{\mathrm{st}}\in E_{\mathrm{st}},
    \quad
    \Delta_{\mathrm{g}}\in E_{\mathrm{g}};
  \]
  \item on $E_{\mathrm{st}}$, the second variation of the population risk is
  coercive:
  \[
    \delta^2 \mc{R}(\rho_T)[h,h]
    \ge
    m_T \norm{h}_{\mathrm{st}}^2
    \qquad \forall h\in E_{\mathrm{st}}
  \]
  for some $m_T>0$;
  \item on $E_{\mathrm{g}}$, the predictive map is flat to first order and the
  risk curvature is much smaller:
  \[
    D\Pi(\rho_T)[h]=0
    \qquad\text{and}\qquad
    \delta^2 \mc{R}(\rho_T)[h,h]
    \le
    \kappa_T \norm{h}_{\mathrm{g}}^2
    \qquad \forall h\in E_{\mathrm{g}},
  \]
  with $\kappa_T\ll m_T$;
  \item after alignment, the two-run finite-width discrepancy satisfies
  \[
    \Delta_N
    :=
    \Theta_A^N-\Theta_B^N
    =
    \Delta_{N,\mathrm{st}}+\Delta_{N,\mathrm{g}},
  \]
  where
  \[
    \norm{\Delta_{N,\mathrm{st}}}=O_{\Prob}(\varepsilon_N),
    \qquad
    \norm{\Delta_{N,\mathrm{g}}}=O_{\Prob}(\gamma_N),
    \qquad
    \varepsilon_N=o(\gamma_N);
  \]
  \item the barrier admits a local lower bound of the form
  \[
    \mathrm{Bar}_N
    \ge
    c_T \norm{\Delta_{N,\mathrm{g}}}^2
    -
    C_T \norm{\Delta_{N,\mathrm{st}}}^2
    +
    o_{\Prob}\!\big(\norm{\Delta_N}^2\big)
  \]
  for some constants $c_T,C_T>0$.
\end{enumerate}
\end{assumption}

\begin{conjecture}[Gauge-dominant finite-width failure for plain MDNs]
\label{conj:gauge-dominance}
Assume the setting of Corollary~\ref{cor:mdn-sgd-common-path} together with
Assumption~\ref{ass:gauge-dominance}. Then the finite-width synthesis barrier
for plain MDNs is asymptotically governed by the gauge component:
\[
  \mathrm{Bar}_N
  =
  c_T \norm{\Delta_{N,\mathrm{g}}}^2
  +
  o_{\Prob}\!\big(\norm{\Delta_{N,\mathrm{g}}}^2\big).
\]
In particular, if the gauge block is nondegenerate in probability, then there
exist constants $\delta,p_0>0$ such that
\[
  \liminf_{N\to\infty}
  \Prob\!\left(
    \mathrm{Bar}_N \ge \delta\,\gamma_N^2
  \right)
  \ge
  p_0.
\]
Thus, once the ordinary mean-field tracking error has been separated out, the
observed run-to-run failure of linear synthesis is explained by gauge
fluctuations rather than by the transverse finite-width error.
\end{conjecture}

\begin{remark}[Why this distinguishes $K=1$ from $K>1$]
The conjecture above is genuinely about the mixture case. For a single Gaussian
head, the nontrivial gauge space is absent or much smaller, so the
$\Delta_{N,\mathrm{g}}$ term disappears and the barrier is controlled by the
ordinary finite-width error alone. The MDN difficulty is therefore not merely
that finite-width approximation is imperfect in both cases; it is that
$K>1$ introduces additional low-curvature directions that can dominate the
observed synthesis failure.
\end{remark}

\begin{remark}[What remains to be proved]
To turn Conjecture~\ref{conj:gauge-dominance} into a theorem, three missing
ingredients have to be supplied.
\begin{enumerate}[label=(\alph*)]
  \item A local quotient geometry for plain MDNs, identifying the stable block
  $E_{\mathrm{st}}$ and the gauge block $E_{\mathrm{g}}$ near the deterministic
  path.
  \item A finite-width fluctuation theorem, after alignment, that shows the
  stable block is smaller than the gauge block in the regime observed
  experimentally.
  \item A lower bound proving that the interpolation barrier is sensitive to the
  gauge block exactly in the sense postulated in (GD5).
\end{enumerate}
These are not generic mean-field tasks. They are precisely the new ingredients
needed to explain why the mixture case fails while the single-Gaussian case does
not.
\end{remark}

\begin{thebibliography}{99}

\bibitem{Bishop2006}
Christopher M. Bishop.
\newblock \emph{Pattern Recognition and Machine Learning}.
\newblock Springer, 2006.

\bibitem{FerbachEtAl2024}
Damien Ferbach, Baptiste Goujaud, Gauthier Gidel, and Aymeric Dieuleveut.
\newblock Proving linear mode connectivity of neural networks via optimal
transport.
\newblock arXiv:2310.19103, 2024.

\bibitem{MeiMisiakiewiczMontanari2019}
Song Mei, Theodor Misiakiewicz, and Andrea Montanari.
\newblock Mean-field theory of two-layers neural networks: Dimension-free
bounds and kernel limit.
\newblock In \emph{Proceedings of the Thirty-Second Conference on Learning
Theory}, pages 2388--2464, 2019.

\end{thebibliography}

\end{document}
