\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{bm}
\usepackage[dvipsnames]{xcolor}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!50!black,
  citecolor=blue!50!black,
  urlcolor=blue!50!black,
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{assumption}[theorem]{Assumption}

\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\PP}{\mathcal{P}}
\newcommand{\W}{W_1}
\newcommand{\dd}{\,\mathrm{d}}

\title{Mean Field Theory for BatchEnsembles}
\author{}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
BatchEnsemble introduces diversity through multiplicative fast weights, but not all such modulations play the same role when width grows. This note studies a fixed-$K$, width-to-infinity mean-field limit for a vector-valued bias-free two-layer BatchEnsemble-style model with neuron-wise fast weights and optional boundary fast weights. The main conclusion is structural and dynamical. In the symmetric regime studied here, width self-averaging does not preserve leading-order head asymmetry coming only from neuron-wise modulation, whereas boundary fast weights survive as finite-dimensional state variables. However, merely making boundary fast weights trainable is not enough to preserve ensemble diversity: under jointly permutation-symmetric initialization of the neuron law and boundary state, the infinite-width predictors still collapse exactly. Even there, collapse means convergence to a common predictor, not automatic equivalence with plain single-model training: in the boundary-only specialization with the head-summed training objective, the collapsed reduction uses the same boundary drift formula as the matched single model at a fixed shared state, while the shared neuron law is accelerated by a factor $K$, so the coupled trajectory generally differs. Within that symmetric regime, the natural route to order-one head asymmetry in the wide feature-learning limit is to seed asymmetry in the finite-dimensional boundary state. We then identify a clean positive theory in a boundary-only subfamily. There, once the shared neuron law is fixed, all heads evolve under the same nonautonomous boundary vector field, so distinct boundary trajectories cannot coalesce in finite time. Moreover, at any fixed time and near the all-ones collapsed boundary point, the leading-order predictor covariance created by small random boundary perturbations for a frozen neuron law is governed by the covariance of that perturbation law filtered through the linearized predictor map. As a result, in the collapsed regime one should not expect leading-order ensemble gain from prediction averaging or leading-order disagreement-based spread, while in this local symmetry-broken regime the useful part of boundary randomness is precisely its projection onto the active sensitivity directions of the predictor map. We prove the rigorous backbone for a controlled clipped version of the concrete model, and then record a finite-horizon continuation principle showing how those clipped conclusions transfer back to the unclipped concrete dynamics under the same bounded-input and bounded-Lipschitz activation hypotheses whenever an a priori bound keeps the trajectory inside an identity region.
\end{abstract}

\section{Introduction}

BatchEnsemble is a parameter-sharing ensemble architecture: instead of training
$K$ fully independent networks, one keeps a shared backbone and gives each head
a small set of multiplicative fast weights. The practical promise is clear. One
would like to obtain ensemble-like diversity without paying the full cost of $K$
separate models. The theoretical question is less clear: \emph{which
head-specific parameters still matter once width becomes very large?}

This note studies that question in a fixed-$K$, width-$M\to\infty$ mean-field
regime. The rigorous backbone is built first for a controlled clipped version
of the concrete model, together with a finite-horizon continuation principle
back to the unclipped concrete dynamics under the same bounded-input and
bounded-Lipschitz activation hypotheses whenever an a priori bound keeps the
trajectory inside an identity region. The key point is
that BatchEnsemble mixes two different kinds of
head-specific structure:
\begin{enumerate}[leftmargin=2em]
  \item \emph{Neuron-wise fast weights}: variables such as
  $s_{\alpha,m}^{(2)}$ and $r_{\alpha,m}^{(1)}$ that are attached to individual
  neurons and therefore, in the symmetric limiting regime studied below,
  self-average as $M\to\infty$.
  \item \emph{Boundary fast weights}: variables such as $s_\alpha^{(1)}$ and
  $r_\alpha^{(2)}$ that sit at the input or output boundary, stay
  finite-dimensional when $M\to\infty$, and therefore do not self-average.
\end{enumerate}
The distinction is operational. Width averaging acts directly on neuron-wise
modulations, but it does not remove asymmetry carried by boundary fast weights.
In the concrete two-layer model introduced in \Cref{sec:prelim}, this leads to
a sharp split between two questions. The architecture decides the macroscopic
state equation: neuron-wise modulations are absorbed into the neuron law,
whereas boundary fast weights produce an additional finite-dimensional boundary
state. Initialization symmetry then decides whether the resulting wide-limit
dynamics collapse across heads: jointly permutation-symmetric initialization of
the neuron law and boundary state yields collapse, while asymmetric boundary
initialization is the natural route to order-one head asymmetry in the
finite-dimensional part of the limit.

The main message is practitioner-facing. If head diversity is implemented only
through neuron-wise modulations, then in the symmetric limiting regime analyzed
here width averaging does not preserve order-one headwise functional separation
at leading order. More strongly, even if boundary fast weights are present and trainable,
jointly permutation-symmetric initialization still leads to a collapsed
infinite-width predictor. If one works in that symmetric neuron-law regime and
wants order-one head asymmetry to survive width averaging, then it must enter
through the boundary state from the start. In the full model, the directly input- and output-connected factors
$s_\alpha^{(1)}$ and $r_\alpha^{(2)}$ play that role. Once such a factor is
retained with head-dependent initialization, the correct limit is no longer a
single PDE but a coupled PDE-ODE system. This
point is especially notable when compared with NTK-based analyses of embedded
ensembles, where one can obtain an \emph{independent regime} under centered
last-layer post-activation modulation \cite{velikanov2022}. Conceptually, this
means that collapse in shared-weight ensembles is not a tautological
consequence of parameter sharing or of head-exchangeable random initialization
laws. The same broad architecture family can display independent infinite-width
behavior in the lazy regime and collapsed leading-order behavior in the
feature-learning regime studied here. Our claim is therefore regime-specific:
in the present feature-learning mean-field limit, neuron-wise modulations alone
do not generically produce leading-order functional independence, and trainable
boundary fast weights do not by themselves prevent collapse under jointly
permutation-symmetric initialization of the mean-field state.

There is also an important nuance inside that collapse statement. A collapsed
wide limit means that all heads realize the same predictor, but it does
\emph{not} automatically mean that the common predictor is generated by the
plain gradient flow of a matched single model. In the boundary-only
specialization studied below, the boundary-side drift has the same functional
form as in a single model when evaluated at the same pair $(\rho_t,q_t)$, while
the shared neuron law is accelerated by a factor $K$. Because that acceleration
changes $\rho_t$, the resulting coupled trajectory is generally different from
ordinary single-model training. So the correct collapsed object is a reduced
common-predictor dynamics that still remembers the shared/private
parameterization.

This matters because practitioners usually do not care about diversity as an
end in itself. They care about what diversity can buy: lower risk after
averaging predictions, better calibration, and useful uncertainty signals from
head disagreement. In the present setting, those benefits require order-one
functional separation across heads. Under exact wide-limit collapse, convex-loss
averaging gains disappear at leading order, and uncertainty proxies built from
head-to-head disagreement vanish at leading order as well. Any such benefit that
remains in the symmetric regime must therefore come from finite-width
fluctuations, optimization asymmetry, stochasticity, or explicit symmetry
breaking in the boundary state.

The negative message is only half the story. Once asymmetry is deliberately
introduced through the boundary state, one would like to know whether training
can immediately wash it out, and, more importantly, which asymmetric directions
are actually useful. The note therefore takes one further step inside a
boundary-only subfamily that keeps only the directly input- and output-connected
fast weights. In that regime, the heads are transported by a common
nonautonomous boundary flow, which already yields a rigorous non-coalescence
statement at the state level. More importantly, at any fixed time and near the
all-ones collapsed boundary point, the leading-order predictor covariance
generated by small random boundary perturbations for a fixed neuron law is
determined by the covariance of the perturbation law projected through the
linearized predictor map. This turns the collapse theorem into a
practitioner-facing local design principle at that fixed-time reference point:
it is not enough to break symmetry; one should allocate boundary variance along
the active sensitivity directions of the wide predictor there.

The rest of the note is organized as follows. \Cref{sec:prelim} introduces the
concrete two-layer model, the neuron-wise versus boundary-fast-weight split, and
a boundary-only specialization used for the positive design theorem.
\Cref{sec:related} positions the paper relative to BatchEnsemble, NTK-based
embedded-ensemble theory, feature-learning mean-field theory, and the tabular
model TabM. \Cref{sec:results} then states the main practitioner-facing
implications for BatchEnsemble design. The abstract mean-field setup, the
general fixed-$K$ width-limit theorem, and all proofs are deferred to the
appendix.

\section{Related Work}
\label{sec:related}

\paragraph{BatchEnsemble and parameter-efficient ensembling.}
BatchEnsemble was introduced by Wen et al.\ \cite{wen2020} as a
parameter-sharing alternative to deep ensembles, in which each ensemble member
is represented by rank-one multiplicative fast weights applied to shared weight
matrices. Our concrete two-layer model is a mean-field idealization of this
construction. The present note is not about improving the architecture itself,
but about understanding which parts of this parameter-efficient ensemble remain
visible in the wide feature-learning limit.

\paragraph{NTK-based operating regimes for embedded ensembles.}
The closest theoretical comparison is the NTK-based analysis of Embedded
Ensembles by Velikanov et al.\ \cite{velikanov2022}. In their lazy-training
setting, infinite-width dynamics are described by a deterministic ensemble NTK,
and centered last-layer post-activation modulation can produce an
\emph{independent regime} in which the off-diagonal GP covariances and
off-diagonal NTK entries vanish. Thus, even with a shared backbone and an
exchangeable iid law for head-wise modulations, the NTK limit need not
collapse. This makes the comparison conceptually important: parameter sharing
and symmetric random initialization do not by themselves force shared-weight
ensembles into a common-predictor limit. Our main message differs precisely at
this point: in the feature-learning mean-field regime studied here, neuron-wise
modulation alone does not generically create leading-order functional
independence, and even trainable boundary fast weights do not prevent collapse
under jointly permutation-symmetric initialization of the mean-field state.

\paragraph{Feature-learning mean-field theory.}
On the mean-field side, two-layer networks admit a deterministic
distributional description in terms of transport PDEs or Wasserstein gradient
flows; see, for example, Mei et al.\ \cite{mei2018} and the lazy-training
perspective of Chizat et al.\ \cite{chizat2019}. Our contribution is not a new
generic mean-field theorem for ordinary networks, but rather an application of
the feature-learning mean-field viewpoint to parameter-efficient ensembles with
shared weights. The outcome is a structural distinction between neuron-wise
fast weights, which self-average with width, and boundary fast weights, which
remain as finite-dimensional state variables.

\paragraph{TabM and the tabular case for efficient ensembling.}
This distinction matters in practice because BatchEnsemble-style ensembling has
already become influential in tabular deep learning. TabM, introduced by
Gorishniy et al.\ \cite{gorishniy2025}, is an MLP-based tabular model built
around parameter-efficient ensembling and reports a strong
performance-efficiency trade-off on public benchmarks. Our results do not
analyze TabM directly, but they help articulate why boundary-connected
modulation mechanisms can be qualitatively different from purely neuron-wise
ones in the feature-learning regime. In that sense, the present theory is
meant as a step toward understanding why parameter-efficient ensembling can be
so effective in domains such as tabular learning.

\section{Preliminary}
\label{sec:prelim}

Fix an input dimension $d_x$, an output dimension $d_y$, a hidden width $M$, and an ensemble size
$K\in\{1,2,\dots\}$. For each neuron $m\in\{1,\dots,M\}$, let
\[
  a_m\in\R^{d_y},\qquad
  w_m\in\R^{d_x},\qquad
  s_{\alpha,m}^{(2)}\in\R\quad (\alpha=1,\dots,K),\qquad
  r_{\alpha,m}^{(1)}\in\R\quad (\alpha=1,\dots,K),
\]
and let
\[
  s^{(1)}=(s_1^{(1)},\dots,s_K^{(1)})\in(\R^{d_x})^K,
  \qquad
  r^{(2)}=(r_1^{(2)},\dots,r_K^{(2)})\in(\R^{d_y})^K.
\]
The two layers take the standard BatchEnsemble form
\[
  W_\alpha = W\circ \bigl(s_\alpha^{(1)}(r_\alpha^{(1)})^\top\bigr),
  \qquad
  A_\alpha = A\circ \bigl(s_\alpha^{(2)}(r_\alpha^{(2)})^\top\bigr),
\]
and the corresponding bias-free two-layer predictor is
\begin{equation}
  f_{\alpha,\mathrm{full}}^M(x)
  :=
  \frac1M A_\alpha^\top \sigma(W_\alpha^\top x).
  \label{eq:model-full-matrix}
\end{equation}
Equivalently, expanding row by row,
\begin{equation}
  f_{\alpha,\mathrm{full}}^M(x)
  :=
  \frac1M\sum_{m=1}^M
  (a_m\odot r_\alpha^{(2)})\,s_{\alpha,m}^{(2)}\,
  \sigma\!\bigl(
    r_{\alpha,m}^{(1)}\,w_m^\top(s_\alpha^{(1)}\odot x)
  \bigr).
  \label{eq:model-full}
\end{equation}
The scalar output case is recovered by taking $d_y=1$.
The empirical-state representation used in the mean-field analysis is deferred
to Appendix~\ref{sec:appendix-setup}, so that the main text can focus on the
model itself and the resulting design implications.

\paragraph{Boundary-only specialization.}
To isolate the role of the directly input- and output-connected fast weights, we
also use the boundary-only subfamily
\begin{equation}
  f_{\alpha,\mathrm{bdry}}^M(x)
  :=
  \frac1M\sum_{m=1}^M
  (a_m\odot v_\alpha)\,
  \sigma\!\bigl(
    w_m^\top(u_\alpha\odot x)
  \bigr),
  \label{eq:model-bdry}
\end{equation}
where
\[
  u_\alpha\in\R^{d_x},
  \qquad
  v_\alpha\in\R^{d_y},
  \qquad
  \theta_m=(a_m,w_m)\in\R^{d_y}\times\R^{d_x}.
\]
This removes the neuron-wise head coordinates and keeps only the boundary state
$h_\alpha=(u_\alpha,v_\alpha)$. After taking the width limit, the corresponding
predictor map at neuron law $\rho$ becomes
\begin{equation}
  H(x;\rho,u,v)
  :=
  \int_{\R^{d_y}\times\R^{d_x}}
  (a\odot v)\,
  \sigma\!\bigl(w^\top(u\odot x)\bigr)\,\rho(\dd a,\dd w).
  \label{eq:H-bdry}
\end{equation}
This is the cleanest setting in which to formulate a positive theorem about how
boundary asymmetry survives width averaging.

\section{Theoretical Results}
\label{sec:results}

\paragraph{Mean-field setup in one sentence.}
The technical results below are stated in terms of the abstract mean-field
objects introduced in Appendix~\ref{sec:appendix-setup}. In that notation, the
large-width limit is described by a neuron law $\rho_t$ and, when boundary fast
weights are present, a finite-dimensional boundary state $h_t$; the full
feature-learning limit is a coupled PDE-ODE system \eqref{eq:pde}--\eqref{eq:ode},
while the no-boundary regime removes the ODE and leaves a single transport PDE.

\paragraph{Guiding question.}
The central design question is which head-specific parameters survive width
averaging. The results below show a sharp split. Neuron-wise modulations
$(s_{\alpha,m}^{(2)},r_{\alpha,m}^{(1)})$ are averaged into the neuron law, while
boundary fast weights remain finite-dimensional state variables. In the full
two-layer model these are $s_\alpha^{(1)}$ and $r_\alpha^{(2)}$. This
structural split is what drives the different practical interpretations of the
two regimes.

\paragraph{Technical note.}
The practitioner-facing statements below are derived from an abstract fixed-$K$
width-limit theorem for the coupled particle/boundary dynamics, together with
the permutation-equivariance structure of the model. To keep the main text
focused on the design message, we defer that general theorem and its
random-initialization corollary to Appendix~\ref{sec:appendix-general-limit},
and defer all proofs to Appendix~\ref{sec:appendix-proofs}.

\subsection{Permutation-Symmetric Initialization Still Collapses}
\label{sec:regime-a}

The first practical question is whether simply adding trainable boundary fast
weights is enough to avoid collapse. The answer is no. This subsection should
be read as covering both the literal no-boundary specialization
\[
  s_\alpha^{(1)}\equiv \mathbf 1,
  \qquad
  r_\alpha^{(2)}\equiv \mathbf 1
  \qquad
  \forall \alpha,
\]
and the full trainable-boundary model initialized on the same permutation-
symmetric diagonal. This is also the bookkeeping used in the NTK analysis of
embedded ensembles: absent modulations are represented by deterministic
variables equal to $1$ almost surely \cite{velikanov2022}. If the neuron law
and boundary state are initialized jointly permutation-symmetrically, then
permutation-equivariant feature-learning dynamics preserve that diagonal
manifold, so the wide-limit predictors still collapse.

\begin{theorem}[Jointly permutation-symmetric initialization yields wide-limit collapse]
\label{thm:identical-boundary-collapse}
Assume \cref{ass:main}, and let $(\rho_t,h_t)$ be the characteristic solution of
\cref{eq:pde,eq:ode} with initial data $(\rho_0,h_0)$. If
\[
  (\Pi_\pi^\Theta)_\#\rho_0=\rho_0,
  \qquad
  \Pi_\pi^\mathcal H h_0=h_0
  \qquad
  \forall \pi\in S_K,
\]
then
\[
  (\Pi_\pi^\Theta)_\#\rho_t=\rho_t,
  \qquad
  \Pi_\pi^\mathcal H h_t=h_t
  \qquad
  \forall t\ge 0,\ \forall \pi\in S_K,
\]
and consequently
\[
  f_1(x;\rho_t,h_t)=\cdots=f_K(x;\rho_t,h_t)
  \qquad
  \forall x\in\mathcal X,\ \forall t\ge 0.
\]
In the concrete BatchEnsemble model
\eqref{eq:model-full-matrix}--\eqref{eq:model-full}, the condition
$\Pi_\pi^\mathcal H h_0=h_0$ means that all heads start from the same
input-side boundary vector $s_\alpha^{(1)}$ and the same output-side boundary
vector $r_\alpha^{(2)}$. The no-boundary model is the special case in which
this common boundary value is fixed to the all-ones vectors.
\end{theorem}

In particular, for the concrete two-layer BatchEnsemble model, if the empirical
neuron law converges to a permutation-symmetric limit and the boundary vectors
$s^{(1)}$ and $r^{(2)}$ are initialized identically across heads, then the
infinite-width predictors collapse exactly.

So \cref{thm:identical-boundary-collapse} should be read as covering two
closely related cases: the standard no-boundary architecture obtained by fixing
$s^{(1)}$ and $r^{(2)}$ to the all-ones vectors, and the full architecture in
which those same boundary coordinates are initialized identically and then
trained. This parallels the NTK treatment of \cite{velikanov2022}, where
removing a modulation is encoded by setting it deterministically to $1$ almost
surely, and the all-ones choice is treated as the fully collective extreme.

Conceptually, this result combines two ingredients: the appendix width-limit
theorem provides the well-posed deterministic limit, and the appendix proof
shows that identical boundary initialization keeps the limit trajectory on the
permutation-symmetric diagonal manifold.

Combined with \cref{cor:random}, this shows that whenever the finite-width
initial environments $(\rho_0^M,h_0^M)$ converge to a deterministic
permutation-symmetric limit $(\rho_0,h_0)$, the infinite-width predictor vector
is deterministic and collapsed across heads.

\begin{remark}[Contrast with NTK-based operating regimes]
\label{rem:ntk-contrast}
This is the sharpest point of contrast with NTK-based analyses of embedded
ensembles. In the lazy-training picture of \cite{velikanov2022}, centered
neuron-wise post-activation modulation can produce an \emph{independent regime}
in which off-diagonal GP covariances and NTK entries vanish. So collapse is not
an architectural tautology of shared-weight ensembling. In that NTK regime, the
heads still share a common backbone, and the head-specific modulations are
drawn from an exchangeable iid law, yet the infinite-width dynamics can be
model-diagonal. By contrast, the present feature-learning mean-field limit
absorbs neuron-wise modulations into the self-averaging neuron law while
retaining only boundary asymmetry as an order-one state variable. Under jointly
permutation-symmetric initialization of that mean-field state, the trajectory
stays on the diagonal and the predictors collapse. Thus NTK-style independence
does not arise here as a law-of-large-numbers effect of neuron-wise modulation
alone; if it appears at all, it belongs to a lazy or fluctuation-level
description rather than to the leading-order feature-learning limit.
\end{remark}

For a practitioner, the message is simple: adding trainable boundary fast
weights is not enough under jointly permutation-symmetric initialization. If
the boundary state is initialized identically across heads and the neuron law is
permutation-symmetric, the wide-limit theory still predicts a
common-predictor leading-order behavior. In the boundary-only specialization,
\cref{prop:boundary-single-flow,cor:boundary-single-kernel} sharpen this
further: the collapsed object is generally not the plain single-model gradient
flow, but a $K$-accelerated-shared version of it. Any head-to-head diversity that
remains at finite width is therefore a subleading fluctuation effect rather than
an order-one property of the feature-learning limit.

\subsection{Randomness Comes from Boundary-State Initialization}
\label{sec:regime-b}

What can be stated in full generality without further nondegeneracy assumptions
is a bookkeeping randomness statement rather than a full non-collapse theorem. Width does
not average away the finite-dimensional boundary state, so in the regime of a
deterministic neuron-law limit, any surviving order-one randomness comes from
the boundary-state initialization rather than from finite-width sampling noise.

\begin{corollary}[Boundary-state initialization is the surviving source of order-one randomness]
\label{cor:boundary-init}
Assume \cref{ass:main}. If $(\rho_0,h_0)$ is deterministic, then the
infinite-width predictor vector is deterministic. If $\rho_0$ is deterministic
but $h_0$ is random, then any order-one randomness that survives at infinite
width is induced entirely by the law of $h_0$.
\end{corollary}

The contrast with \cref{thm:identical-boundary-collapse} is therefore not the
mere presence of boundary fast weights, but the fact that width does not erase
the finite-dimensional boundary state. A stronger theorem asserting that every
asymmetric boundary initialization remains functionally separated for all times
would require additional nondegeneracy assumptions on the boundary dynamics and
on the predictor map $h\mapsto f(\cdot;\rho_t,h)$. In particular,
\cref{cor:boundary-init} by itself does not claim persistent disagreement or
visible non-collapse; it only identifies where order-one randomness can enter
the deterministic wide limit.

\subsection{A Positive Boundary-Only Theory After Symmetry Breaking}
\label{sec:boundary-positive}

\Cref{cor:boundary-init} identifies the source of order-one randomness, but it
does not yet say when that randomness remains functionally visible. The
boundary-only specialization \eqref{eq:model-bdry} is the cleanest regime in
which one can go further without making a false global non-collapse claim.

\begin{proposition}[Boundary-only heads are transported by a common vector field]
\label{prop:boundary-flow}
Consider the boundary-only model \eqref{eq:model-bdry}, and let $\rho_t$ denote
its fixed-$K$ width-limit neuron law. For
\(
  q=(u,v)\in\R^{d_x}\times\R^{d_y}
\),
define
\begin{equation}
  C(\rho,q)
  :=
  -\E_{(x,y)\sim\mathsf P}
  \Bigl[
    \int_{\R^{d_y}\times\R^{d_x}}
    \nabla_q\!\Bigl(
      (a\odot v)\,\sigma\!\bigl(w^\top(u\odot x)\bigr)
    \Bigr)^\top
    \nabla_1\ell(H(x;\rho,u,v),y)\,\rho(\dd a,\dd w)
  \Bigr].
  \label{eq:boundary-drift}
\end{equation}
Then the boundary states satisfy
\[
  \dot h_{\alpha,t}=C(\rho_t,h_{\alpha,t}),
  \qquad
  \alpha=1,\dots,K.
\]
If, in addition,
\[
  \|C(\rho_t,q)-C(\rho_t,\bar q)\|
  \le
  L_t\|q-\bar q\|
  \qquad
  \forall q,\bar q\in\R^{d_x}\times\R^{d_y},
\]
for some integrable rate $L_t$ on $[0,T]$, and if
\[
  \Lambda_t:=\int_0^t L_s\,\dd s,
\]
let $\Phi_{s,t}$ denote the associated flow maps on $[0,T]$, so that
uniqueness gives $\Phi_{t,s}=(\Phi_{s,t})^{-1}$. Then for any two heads
$\alpha,\beta$ and all $t\in[0,T]$,
\begin{equation}
  e^{-\Lambda_t}\,
  \|h_{\alpha,0}-h_{\beta,0}\|
  \le
  \|h_{\alpha,t}-h_{\beta,t}\|
  \le
  e^{\Lambda_t}\,
  \|h_{\alpha,0}-h_{\beta,0}\|.
  \label{eq:boundary-bilipschitz}
\end{equation}
In particular, identical boundary initialization stays identical, while
distinct boundary states cannot coalesce in finite time.
\end{proposition}

\begin{proposition}[Collapsed boundary-only flow is a shared-state reduction with $K$-accelerated neuron law]
\label{prop:boundary-single-flow}
Consider the boundary-only model \eqref{eq:model-bdry}, and define the
corresponding single-model population risk
\[
  \mathcal R_{\mathrm{single}}(\rho,q)
  :=
  \E_{(x,y)\sim\mathsf P}
  \bigl[\ell(H(x;\rho,q),y)\bigr],
  \qquad
  q=(u,v)\in\R^{d_x}\times\R^{d_y}.
\]
Define the associated single-model drifts
\begin{align}
  B_{\mathrm{single}}((a,w);\rho,q)
  &:=
  -\E_{(x,y)\sim\mathsf P}
  \Bigl[
    \nabla_{(a,w)}\!\Bigl(
      (a\odot v)\,\sigma\!\bigl(w^\top(u\odot x)\bigr)
    \Bigr)^\top
    \nabla_1\ell(H(x;\rho,q),y)
  \Bigr],
  \label{eq:single-bdry-B}\\
  C_{\mathrm{single}}(\rho,q)
  &:=
  -\E_{(x,y)\sim\mathsf P}
  \Bigl[
    \int_{\R^{d_y}\times\R^{d_x}}
    \nabla_q\!\Bigl(
      (a\odot v)\,\sigma\!\bigl(w^\top(u\odot x)\bigr)
    \Bigr)^\top
    \nabla_1\ell(H(x;\rho,q),y)\,\rho(\dd a,\dd w)
  \Bigr].
  \label{eq:single-bdry-C}
\end{align}
Assume the fixed-$K$ width-limit trajectory is initialized on the diagonal,
\[
  h_{1,0}=\cdots=h_{K,0}=:q_0.
\]
Then
\[
  h_{1,t}=\cdots=h_{K,t}=:q_t
  \qquad
  \forall t\ge 0,
\]
and the pair $(\rho_t,q_t)$ solves
\begin{align}
  \partial_t\rho_t
  +
  \nabla_{(a,w)}\cdot
  \bigl(
    \rho_t\,K\,B_{\mathrm{single}}((a,w);\rho_t,q_t)
  \bigr)
  &=0,
  \label{eq:diag-single-pde}\\
  \dot q_t
  &=
  C_{\mathrm{single}}(\rho_t,q_t).
  \label{eq:diag-single-ode}
\end{align}
Thus the collapsed boundary-only BatchEnsemble limit uses the same
boundary-drift functional form as ordinary single-model mean-field training
when evaluated at the same pair $(\rho_t,q_t)$, but its shared neuron law moves
with an additional factor $K$.
\end{proposition}

\begin{corollary}[Collapsed boundary-only predictor retains the shared/private speed split]
\label{cor:boundary-single-kernel}
Under the assumptions of \cref{prop:boundary-single-flow}, assume in addition
that $t\mapsto(\rho_t,q_t)$ is regular enough that differentiation under the
integral sign is valid. Then the common predictor
\[
  f(x,t):=H(x;\rho_t,q_t)
\]
satisfies
\begin{equation}
  \partial_t f(x,t)
  =
  -\E_{(x',y)\sim\mathsf P}
  \Bigl[
    \Gamma_{\mathrm{eff}}^{\mathrm{bdry}}(x,x';t)\,
    \nabla_1\ell(f(x',t),y)
  \Bigr],
  \label{eq:collapsed-bdry-output}
\end{equation}
with effective kernel
\begin{equation}
  \Gamma_{\mathrm{eff}}^{\mathrm{bdry}}(x,x';t)
  :=
  K\,\Gamma_{\mathrm{neu}}^{\mathrm{bdry}}(x,x';t)
  +
  \Gamma_{\mathrm{bdry}}^{\mathrm{bdry}}(x,x';t),
  \label{eq:collapsed-bdry-keff}
\end{equation}
where
\begin{align}
  \Gamma_{\mathrm{neu}}^{\mathrm{bdry}}(x,x';t)
  &:=
  \int_{\R^{d_y}\times\R^{d_x}}
  \nabla_{(a,w)}\widehat\Psi(x;(a,w),q_t)^\top
  \nabla_{(a,w)}\widehat\Psi(x';(a,w),q_t)
  \rho_t(\dd a,\dd w),
  \label{eq:collapsed-bdry-kneu}\\
  \Gamma_{\mathrm{bdry}}^{\mathrm{bdry}}(x,x';t)
  &:=
  \nabla_q H(x;\rho_t,q_t)^\top
  \nabla_q H(x';\rho_t,q_t),
  \label{eq:collapsed-bdry-kbdry}
\end{align}
and
\[
  \widehat\Psi(x;(a,w),q)
  :=
  (a\odot v)\,\sigma\!\bigl(w^\top(u\odot x)\bigr).
\]
In particular, the plain single-model kernel flow for the same architecture
would use
\(
  \Gamma_{\mathrm{neu}}^{\mathrm{bdry}}+\Gamma_{\mathrm{bdry}}^{\mathrm{bdry}}
\)
instead. So, for $K>1$, the collapsed BatchEnsemble dynamics is generally not
the same as training an ordinary single model with equal learning rates; at
the level of this continuous-time limit it is recovered by multiplying the
shared-side learning rate of the matched single-model baseline by $K$
(equivalently, dividing the shared-side learning rate on the BatchEnsemble
side by $K$), or in
degenerate cases such as
\(
  \Gamma_{\mathrm{neu}}^{\mathrm{bdry}}\equiv 0
\). The boundary-side drift has the same functional form in the two systems,
but the actual coupled boundary trajectory generally differs because it is
driven by a different shared-law path $\rho_t$.
\end{corollary}

\Cref{prop:boundary-flow} is a state-level statement. For the bounded-Lipschitz
class treated rigorously here, upgrading state separation to a global
function-level theorem would require additional injectivity or lower-Lipschitz
assumptions on the predictor map $q\mapsto H(\cdot;\rho_t,q)$, and we do not
impose them. In the raw positively homogeneous bias-free boundary model there
is an additional obstruction, namely the scale gauge
\[
  H(x;\rho,cu,c^{-1}v)=H(x;\rho,u,v)
  \qquad
  \forall c>0.
\]
which is best read as a supplementary warning for the untruncated homogeneous
setting rather than as part of the formal hypothesis class of the theorem
below. The next result is therefore intentionally local and fixed-time: it
asks which small perturbation directions around the all-ones collapsed boundary
point create leading-order predictor spread for a fixed neuron law $\rho_t$.

\begin{theorem}[Fixed-time local covariance expansion around the all-ones collapsed boundary point]
\label{thm:local-covariance}
Fix a time $t\ge 0$, let $\mathsf P_X$ be the marginal of $\mathsf P$ on
$\mathcal X$, and define
\[
  H_t(u,v)
  :=
  H(\cdot;\rho_t,u,v)
  \in
  L^2(\mathsf P_X;\R^{d_y}).
\]
Assume that $H_t$ is $C^2$ on an open neighborhood $\mathcal N$ of
$(\mathbf 1,\mathbf 1)$ and that its second derivative is bounded on
$\mathcal N$. Let
\[
  \xi\in\R^{d_x},
  \qquad
  \zeta\in\R^{d_y}
\]
be centered essentially bounded random vectors such that
\[
  (\mathbf 1+\varepsilon\xi,\mathbf 1+\varepsilon\zeta)\in\mathcal N
  \qquad
  \text{a.s.}
\]
for all sufficiently small $\varepsilon>0$. Define
\[
  U_\varepsilon:=\mathbf 1+\varepsilon\xi,
  \qquad
  V_\varepsilon:=\mathbf 1+\varepsilon\zeta,
  \qquad
  Y_t^\varepsilon:=H_t(U_\varepsilon,V_\varepsilon).
\]
Write
\[
  g_t:=H_t(\mathbf 1,\mathbf 1),
  \qquad
  A_t:=D_u H_t(\mathbf 1,\mathbf 1),
  \qquad
  B_t:=D_v H_t(\mathbf 1,\mathbf 1),
\]
and let
\[
  \Sigma_u:=\E[\xi\xi^\top],
  \qquad
  \Sigma_v:=\E[\zeta\zeta^\top],
  \qquad
  \Sigma_{uv}:=\E[\xi\zeta^\top].
\]
Then:
\begin{enumerate}[label=(\roman*),leftmargin=2em]
  \item there exists a remainder
  \(
    R_t^\varepsilon\in L^2(\mathsf P_X;\R^{d_y})
  \)
  such that
  \begin{equation}
    Y_t^\varepsilon
    =
    g_t+\varepsilon(A_t\xi+B_t\zeta)+R_t^\varepsilon,
    \qquad
    \E\bigl[\|R_t^\varepsilon\|_{L^2(\mathsf P_X)}^2\bigr]
    =
    O(\varepsilon^4);
    \label{eq:local-linearization}
  \end{equation}
  \item the covariance operator of $Y_t^\varepsilon$ satisfies
  \begin{equation}
    \operatorname{Cov}(Y_t^\varepsilon)
    =
    \varepsilon^2\mathcal K_t+o(\varepsilon^2)
    \label{eq:local-cov-op}
  \end{equation}
  in operator norm on $L^2(\mathsf P_X;\R^{d_y})$, where
  \begin{equation}
    \mathcal K_t
    :=
    A_t\Sigma_u A_t^*
    +
    B_t\Sigma_v B_t^*
    +
    A_t\Sigma_{uv} B_t^*
    +
    B_t\Sigma_{uv}^\top A_t^*;
    \label{eq:local-cov-operator}
  \end{equation}
  \item if $\mathcal K_t\neq 0$, then the law of $Y_t^\varepsilon$ is non-Dirac
  for all sufficiently small $\varepsilon>0$.
\end{enumerate}
For the concrete boundary-only model, $A_t$ and $B_t$ act as
\[
  (A_t q)(x)=J_t(x)q,
  \qquad
  (B_t r)(x)=\operatorname{Diag}(g_t(x))\,r,
\]
where
\[
  J_t(x):=D_u H_t(\mathbf 1,\mathbf 1)(x)\in\R^{d_y\times d_x},
\]
so $\mathcal K_t$ is induced by the kernel
\begin{align}
  K_t(x,x')
  &=
  J_t(x)\Sigma_u J_t(x')^\top
  +
  \operatorname{Diag}(g_t(x))\Sigma_v\operatorname{Diag}(g_t(x'))
  \notag\\
  &\quad
  +
  J_t(x)\Sigma_{uv}\operatorname{Diag}(g_t(x'))
  +
  \operatorname{Diag}(g_t(x))\Sigma_{uv}^\top J_t(x')^\top.
  \label{eq:local-cov-kernel}
\end{align}
\end{theorem}

\begin{corollary}[Leading-order disagreement is a covariance-trace]
\label{cor:boundary-local-disagreement}
Under the assumptions of \cref{thm:local-covariance}, let
$Y_t^{\varepsilon,1}$ and $Y_t^{\varepsilon,2}$ be iid copies of
$Y_t^\varepsilon$. Then
\begin{equation}
  \E\bigl[
    \|Y_t^{\varepsilon,1}-Y_t^{\varepsilon,2}\|_{L^2(\mathsf P_X)}^2
  \bigr]
  =
  2\varepsilon^2\operatorname{Tr}(\mathcal K_t)+o(\varepsilon^2).
  \label{eq:local-disagreement-trace}
\end{equation}
Equivalently,
\begin{equation}
  \operatorname{Tr}(\mathcal K_t)
  =
  \operatorname{Tr}(M_t\Sigma),
  \qquad
  M_t
  :=
  \begin{pmatrix}
    A_t^*A_t & A_t^*B_t\\
    B_t^*A_t & B_t^*B_t
  \end{pmatrix},
  \label{eq:local-design-matrix}
\end{equation}
where
\(
  \Sigma=
  \begin{psmallmatrix}
    \Sigma_u & \Sigma_{uv}\\
    \Sigma_{uv}^\top & \Sigma_v
  \end{psmallmatrix}
\)
is the covariance of $(\xi,\zeta)$. In particular, under a fixed trace budget
for positive-semidefinite $\Sigma$, the leading-order disagreement is maximized
by placing boundary covariance along the top eigendirections of $M_t$.
\end{corollary}

The point of \cref{thm:local-covariance,cor:boundary-local-disagreement} is that
they convert the qualitative slogan ``break symmetry in the boundary state''
into a quantitative fixed-time frozen-law design rule. Near the all-ones
collapsed boundary point, not all symmetry-breaking directions matter equally:
only the part of the boundary covariance that is seen through the active
sensitivity operators $A_t$ and $B_t$ contributes at leading order, while
directions in the kernel or near-kernel of the linearized predictor map
contribute little or nothing.

\begin{remark}[The local covariance theorem is a frozen-law statement]
\label{rem:local-frozen-law}
\Cref{thm:local-covariance,cor:boundary-local-disagreement} perturb the
boundary argument of $H(\cdot;\rho_t,u,v)$ at a fixed time with $\rho_t$ held
fixed. They therefore describe the local sensitivity of the predictor map at
that reference law, not the full linear response of the coupled PDE-ODE
solution map from boundary initialization at time $0$ to the predictor at time
$t$. To obtain the latter one would need to linearize the coupled flow itself,
which introduces an additional first-order contribution through the induced
perturbation of $\rho_t$.
\end{remark}

\subsection{Implications for Ensemble Gain and Uncertainty}
\label{sec:gain-uncertainty}

To connect the collapse theorem with practitioner-facing metrics, define the
head average
\[
  \bar f(x;\rho,h)
  :=
  \frac1K\sum_{\alpha=1}^K f_\alpha(x;\rho,h),
\]
and the corresponding ensemble risk
\[
  \mathcal R_{\mathrm{ens}}(\rho,h)
  :=
  \E_{(x,y)\sim\mathsf P}
  \bigl[\ell(\bar f(x;\rho,h),y)\bigr].
\]
Recall that \eqref{eq:risk} is the \emph{sum} of the per-head training risks, so
\[
  \overline{\mathcal R}(\rho,h):=\frac1K\mathcal R(\rho,h)
\]
is their average.

\paragraph{Ensemble gain under convex losses.}
If $\ell(\cdot,y)$ is convex for every $y$, then Jensen's inequality gives
\[
  \mathcal R_{\mathrm{ens}}(\rho,h)\le \overline{\mathcal R}(\rho,h).
\]
Thus averaging can only help through functional disagreement across heads. When
\[
  f_1(\cdot;\rho,h)=\cdots=f_K(\cdot;\rho,h),
\]
one has $\bar f=f_\alpha$ for every $\alpha$, so the Jensen gap is exactly zero:
\[
  \mathcal R_{\mathrm{ens}}(\rho,h)=\overline{\mathcal R}(\rho,h).
\]
Therefore \cref{thm:identical-boundary-collapse} implies that, in the jointly
permutation-symmetric regime, the wide-limit theory predicts no leading-order
ensemble gain from prediction averaging.

For squared loss this generic Jensen statement sharpens to an exact identity.
Writing
\[
  \ell(\widehat y,y)=\|\widehat y-y\|^2,
\]
one has pointwise in $(x,y)$
\[
  \frac1K\sum_{\alpha=1}^K \|f_\alpha(x;\rho,h)-y\|^2
  =
  \|\bar f(x;\rho,h)-y\|^2
  +
  \frac1K\sum_{\alpha=1}^K \|f_\alpha(x;\rho,h)-\bar f(x;\rho,h)\|^2.
\]
The second term is exactly the disagreement energy, so for squared loss the
ensemble gain relative to $\overline{\mathcal R}$ is precisely the reduction in
head-to-head functional variance.
Combined with \cref{cor:boundary-local-disagreement}, this gives the positive
fixed-time frozen-law counterpart to collapse in the boundary-only regime: once
small centered boundary perturbations are introduced near the all-ones
collapsed point, the first nonvanishing disagreement term and hence the first
nonvanishing squared-loss averaging gain for that reference law are both
governed by
$\operatorname{Tr}(\mathcal K_t)$.

\paragraph{Interpretation: uncertainty from head disagreement.}
At a fixed input $x$, define the ensemble spread
\[
  U(x;\rho,h)
  :=
  \frac1K\sum_{\alpha=1}^K
  \|f_\alpha(x;\rho,h)-\bar f(x;\rho,h)\|^2.
\]
If the heads collapse, then $U(x;\rho,h)\equiv 0$. Hence in the collapsed
wide-limit regime, any uncertainty proxy that is built purely from disagreement
across head predictions also vanishes at leading order. This includes the usual
ensemble variance in regression and, at an interpretive level after passing to
predictive probabilities, disagreement-type classification diagnostics such as
variance across members or mutual-information-style epistemic scores.

The qualifier matters: collapse does \emph{not} imply that every notion of
uncertainty vanishes. The common predictor may still have nonzero predictive
entropy or incur nontrivial loss because of aleatoric uncertainty or ordinary
single-model uncertainty. What disappears at leading order is specifically the
ensemble-induced \emph{epistemic spread} coming from differences between heads.

\paragraph{Operational reading.}
In the symmetric wide regime, one should not interpret a BatchEnsemble as a
source of order-one averaging gains or order-one disagreement-based epistemic
uncertainty. If such effects are observed at finite width, then the theory says
they must come from subleading finite-width fluctuations, stochastic training,
optimizer-induced asymmetry, data-order effects, or explicit symmetry breaking
in the boundary-state initialization.

\subsection{Implications for BatchEnsemble design}

Three separate axes matter.

\paragraph{Architecture axis.}
Once trainable boundary fast weights are included, the correct large-width state
is the pair $(\rho_t,h_t)$ rather than the neuron law alone. The special case
with no boundary fast weights is recovered by freezing $h_t$ to the all-ones
state. Correspondingly, architecture decides the state equation: without
trainable boundary fast weights one gets a single transport PDE, whereas with
trainable boundary fast weights one gets a coupled PDE-ODE system. Even on the
collapsed diagonal of the boundary-only regime, \cref{prop:boundary-single-flow,cor:boundary-single-kernel}
show that the resulting common predictor is generally not the plain
single-model flow: the shared neuron law is accelerated relative to the matched
single-model baseline, and although the boundary-side drift has the same
functional form, it is evaluated along a different shared-law trajectory.

\paragraph{Initialization axis.}
Within a given state equation, initialization symmetry decides whether the wide
limit collapses across heads. Jointly permutation-symmetric initialization
forces exact head collapse by \cref{thm:identical-boundary-collapse}. By
\cref{cor:boundary-init}, if the neuron-law limit is deterministic, then
surviving order-one randomness can come only from the initial law of the
finite-dimensional boundary state. This is only a source-identification
statement; whether that randomness becomes functionally visible is a separate
question addressed locally by \cref{thm:local-covariance}.

\paragraph{Local sensitivity axis.}
Once symmetry is broken in the boundary state, not every direction is equally
useful. In the boundary-only local regime of
\cref{thm:local-covariance,cor:boundary-local-disagreement}, the leading-order
functional spread created by small random boundary perturbations is
$2\varepsilon^2\operatorname{Tr}(\mathcal K_t)+o(\varepsilon^2)$, or
equivalently $\operatorname{Tr}(M_t\Sigma)$ in the finite-dimensional block
notation of \cref{eq:local-design-matrix}. Thus the useful part of boundary
randomness is its projection onto the active eigendirections of the linearized
predictor map, while directions in the kernel or near-kernel do not produce
leading-order functional separation.

\paragraph{Operational takeaway.}
If the goal is to preserve order-one head diversity at very large width, then a
boundary fast weight is structurally different from a neuron-wise fast weight,
but merely making it trainable is not enough. In the symmetric neuron-law
regime covered by \cref{thm:identical-boundary-collapse}, the directly
input- and output-connected factors $s^{(1)}$ and $r^{(2)}$ matter at leading
order only when their initialization breaks head symmetry. In particular, under
convex deployment losses, this symmetry breaking is what can create a
leading-order Jensen gain from averaging, and under disagreement-based
uncertainty estimation it is what can locally sustain order-one epistemic spread across
heads. The local covariance theorem sharpens this further at a fixed-time
frozen-law reference point: one should not merely increase boundary variance,
but rather place it along the active sensitivity directions of the wide
predictor there. At the level proved here, the relevant design object is the
covariance of the boundary perturbation at that reference point, not just the
fact that the perturbation is asymmetric; relating this object back to the
boundary initialization law would require linearizing the full coupled training
map as noted in \cref{rem:local-frozen-law}.

\begin{thebibliography}{9}

\bibitem{wen2020}
Yeming Wen, Dustin Tran, and Jimmy Ba.
\newblock BatchEnsemble: An alternative approach to efficient ensemble and
lifelong learning.
\newblock In \emph{International Conference on Learning Representations}, 2020.

\bibitem{jacot2018}
Arthur Jacot, Franck Gabriel, and Cl\'ement Hongler.
\newblock Neural tangent kernel: Convergence and generalization in neural
networks.
\newblock In \emph{Advances in Neural Information Processing Systems}, 2018.

\bibitem{lee2019}
Jaehoon Lee, Lechao Xiao, Samuel S. Schoenholz, Yasaman Bahri, Jascha
Sohl-Dickstein, and Jeffrey Pennington.
\newblock Wide neural networks of any depth evolve as linear models under
gradient descent.
\newblock In \emph{Advances in Neural Information Processing Systems}, 2019.

\bibitem{chizat2019}
Lena\"ic Chizat, Edouard Oyallon, and Francis Bach.
\newblock On lazy training in differentiable programming.
\newblock In \emph{Advances in Neural Information Processing Systems}, 2019.

\bibitem{mei2018}
Song Mei, Andrea Montanari, and Phan-Minh Nguyen.
\newblock A mean field view of the landscape of two-layer neural networks.
\newblock \emph{Proceedings of the National Academy of Sciences}, 115(33), 2018.

\bibitem{velikanov2022}
Maksim Velikanov, Roman V. Kail, Ivan Anokhin, Roman Vashurin, Maxim Panov,
Alexey Zaytsev, and Dmitry Yarotsky.
\newblock Embedded Ensembles: Infinite width limit and operating regimes.
\newblock In \emph{Proceedings of The 25th International Conference on
Artificial Intelligence and Statistics}, 2022.

\bibitem{gorishniy2025}
Yury Gorishniy, Akim Kotelnikov, and Artem Babenko.
\newblock TabM: Advancing tabular deep learning with parameter-efficient
ensembling.
\newblock In \emph{International Conference on Learning Representations}, 2025.

\end{thebibliography}

\appendix

\section{Abstract Mean-Field Setup and Assumptions}
\label{sec:appendix-setup}

\subsection{Concrete empirical state for the BatchEnsemble model}

For the concrete model \eqref{eq:model-full-matrix}--\eqref{eq:model-full}, the
width-$M$ empirical measure is taken only over the neuron-indexed coordinates:
\begin{equation}
  \rho^M
  :=
  \frac1M\sum_{m=1}^M \delta_{\theta_m},
  \qquad
  \theta_m=
  (a_m,w_m,s_{1,m}^{(2)},\dots,s_{K,m}^{(2)},r_{1,m}^{(1)},\dots,r_{K,m}^{(1)}).
  \label{eq:empirical}
\end{equation}
The boundary fast weights are kept outside this empirical measure:
\begin{equation}
  \bar h^M := (s^{(1)},r^{(2)}) \in (\R^{d_x})^K\times(\R^{d_y})^K.
  \label{eq:boundary-concrete}
\end{equation}
Thus the full finite-width state is the pair $(\rho^M,\bar h^M)$.

\subsection{Abstract mean-field formulation}

At the abstract level we compress all boundary fast weights into a single symbol
$h$. Concretely, for the BatchEnsemble family above this symbol will stand for
the pair $\bar h^M=(s^{(1)},r^{(2)})$. Let $\Theta=\R^D$ denote the
single-neuron state space and let
\[
  \mathcal H=\R^{K d_h}
\]
denote the space of boundary states. Write
\[
  h=(h_1,\dots,h_K)\in\mathcal H,
  \qquad
  h_\alpha\in\R^{d_h}.
\]
Let $(\mathcal X\times\mathcal Y,\mathsf P)$ be the data space.
Fix an output dimension $d_y\in\{1,2,\dots\}$.

For each head $\alpha\in\{1,\dots,K\}$, let
\[
  \Psi_\alpha:\mathcal X\times\Theta\times\mathcal H\to\R^{d_y}
\]
be a feature map. Given a neuron law $\rho\in\PP_1(\Theta)$ and a boundary state
$h\in\mathcal H$, define
\begin{equation}
  f_\alpha(x;\rho,h)
  :=
  \int_\Theta \Psi_\alpha(x;\theta,h)\,\rho(\dd\theta).
  \label{eq:output}
\end{equation}
The head-summed population training risk is
\begin{equation}
  \mathcal R(\rho,h)
  :=
  \E_{(x,y)\sim\mathsf P}
  \Bigl[
    \sum_{\alpha=1}^K \ell(f_\alpha(x;\rho,h),y)
  \Bigr].
  \label{eq:risk}
\end{equation}
This sum-normalized convention keeps the dependence on the number of heads
$K$ visible in the shared update channel. Since $K$ is fixed throughout, one
could divide \eqref{eq:risk} by $K$ without changing its minimizers, but that
would move the explicit factor of $K$ from the shared drift to the head-local
drift in the collapsed single-model comparison below.

\begin{definition}[Mean-field drifts]
Assume $\ell$ is differentiable in its first argument. Define
\begin{equation}
  B(\theta;\rho,h)
  :=
  -\E_{(x,y)\sim\mathsf P}
  \Bigl[
    \sum_{\alpha=1}^K
    \nabla_\theta \Psi_\alpha(x;\theta,h)^\top
    \nabla_1\ell(f_\alpha(x;\rho,h),y)
  \Bigr]
  \label{eq:drift-B}
\end{equation}
and
\begin{equation}
  G_\alpha(\rho,h)
  :=
  -\E_{(x,y)\sim\mathsf P}
  \Bigl[
    \int_\Theta
    \nabla_{h_\alpha}\Psi_\alpha(x;\theta,h)^\top
    \nabla_1\ell(f_\alpha(x;\rho,h),y)\,\rho(\dd\theta)
  \Bigr].
  \label{eq:drift-G}
\end{equation}
Write
\[
  G(\rho,h):=(G_\alpha(\rho,h))_{\alpha=1}^K\in\mathcal H.
\]
The fixed-$K$ mean-field system is
\begin{align}
  \partial_t \rho_t
  +
  \nabla_\theta\cdot\bigl(\rho_t B(\theta;\rho_t,h_t)\bigr)
  &=0,
  \label{eq:pde}\\
  \dot h_t &= G(\rho_t,h_t).
  \label{eq:ode}
\end{align}
\end{definition}

\begin{definition}[Characteristic solutions]
Fix $T>0$ and an initial condition $(\rho_0,h_0)\in\PP_1(\Theta)\times\mathcal H$.
A pair
\[
  (\rho_t,h_t)_{t\in[0,T]}
  \in
  C\bigl([0,T],\PP_1(\Theta)\times\mathcal H\bigr)
\]
is a characteristic solution of \cref{eq:pde,eq:ode} if there exists a flow
\[
  \Phi:[0,T]\times\Theta\to\Theta
\]
such that, for $\rho_0$-almost every $\theta$,
\[
  \dot\Phi_t(\theta)=B(\Phi_t(\theta);\rho_t,h_t),
  \qquad
  \Phi_0(\theta)=\theta,
  \qquad
  \rho_t=(\Phi_t)_\#\rho_0,
\]
while $h_t$ solves \cref{eq:ode}.
\end{definition}

\subsection{Structural assumptions}

Let $S_K$ be the permutation group on $\{1,\dots,K\}$. For $\pi\in S_K$, let
\[
  \Pi_\pi^\mathcal H(h_1,\dots,h_K)
  =
  (h_{\pi^{-1}(1)},\dots,h_{\pi^{-1}(K)}).
\]
Assume also that for each $\pi$ there is a linear isometry
$\Pi_\pi^\Theta:\Theta\to\Theta$ representing the corresponding permutation of
head-labelled coordinates inside the neuron state.

\begin{assumption}[Bounded-Lipschitz structure and permutation equivariance]
\label{ass:main}
There exist finite constants
\[
  B_\Psi,\ B_{\nabla\Psi},\ L_\Psi,\ L_{\nabla\Psi},\ L_\ell,\ C_{\ell,\Psi}
\]
such that the following hold uniformly over all $\alpha\in\{1,\dots,K\}$ and
$x\in\mathcal X$.
\begin{enumerate}[label=(\roman*),leftmargin=2em]
  \item The maps $\Psi_\alpha(x;\cdot,\cdot)$, $\nabla_\theta\Psi_\alpha(x;\cdot,\cdot)$,
  and $\nabla_{h_\alpha}\Psi_\alpha(x;\cdot,\cdot)$ are globally bounded in
  Euclidean/operator norm by $B_\Psi$ and $B_{\nabla\Psi}$. They are also
  globally Lipschitz in $(\theta,h)$ with constants $L_\Psi$ and
  $L_{\nabla\Psi}$.
  \item $\nabla_1\ell(\cdot,y)$ is globally Lipschitz with constant $L_\ell$
  for all $y\in\mathcal Y$, and
  \begin{equation}
    \E_{Y\sim\mathsf P_Y}
    \Bigl[
      \sup_{\|z\|\le B_\Psi}\|\nabla_1\ell(z,Y)\|
    \Bigr]
    \le
    C_{\ell,\Psi},
    \label{eq:loss-envelope}
  \end{equation}
  where $\mathsf P_Y$ denotes the $Y$-marginal of $\mathsf P$.
  \item The feature maps are permutation-equivariant:
  \begin{equation}
    \Psi_\alpha(x;\Pi_\pi^\Theta\theta,\Pi_\pi^\mathcal H h)
    =
    \Psi_{\pi^{-1}(\alpha)}(x;\theta,h)
    \qquad
    \forall \pi\in S_K.
    \label{eq:equiv}
  \end{equation}
\end{enumerate}
\end{assumption}

\begin{remark}[Squared loss now fits the abstract loss hypothesis]
\label{rem:squared-loss}
Because \cref{ass:main}(i) implies
\(
  \|f_\alpha(x;\rho,h)\|\le B_\Psi
\)
for every head, input, and state, the loss assumption only needs control of
$\nabla_1\ell$ on the predictor range actually visited by the clipped model. In
particular, squared loss
\[
  \ell(\widehat y,y)=\|\widehat y-y\|^2
\]
fits \cref{ass:main}(ii) whenever $\E[\|Y\|]<\infty$, since then
\[
  \nabla_1\ell(z,y)=2(z-y),
  \qquad
  L_\ell=2,
  \qquad
  C_{\ell,\Psi}\le 2B_\Psi+2\E[\|Y\|].
\]
\end{remark}

\begin{remark}[Concrete BatchEnsemble family]
\label{rem:concrete}
For the full model \eqref{eq:model-full-matrix}--\eqref{eq:model-full}, take
\[
  \Theta_{\mathrm{BE}}
  =
  \R^{d_y}\times\R^{d_x}\times\R^K\times\R^K,
  \qquad
  \theta=(a,w,s_1^{(2)},\dots,s_K^{(2)},r_1^{(1)},\dots,r_K^{(1)}),
  \qquad
  \mathcal H_{\mathrm{BE}}=(\R^{d_x})^K\times(\R^{d_y})^K.
\]
Write a generic boundary state as
\[
  h=(u,v)\in\mathcal H_{\mathrm{BE}},
  \qquad
  u=(u_1,\dots,u_K)\in(\R^{d_x})^K,
  \qquad
  v=(v_1,\dots,v_K)\in(\R^{d_y})^K.
\]
Then \eqref{eq:model-full} is of the abstract form \eqref{eq:output} with
\begin{equation}
  \Psi_\alpha^{\mathrm{BE}}(x;\theta,h)
  :=
  (a\odot v_\alpha)\,s_\alpha^{(2)}\,
  \sigma\!\bigl(r_\alpha^{(1)}\,w^\top(u_\alpha\odot x)\bigr).
  \label{eq:psi-concrete}
\end{equation}
Fixing $(u_\alpha,v_\alpha)\equiv(\mathbf 1,\mathbf 1)$ for all heads removes
the boundary-state dependence, equivalently making the feature maps independent of
the dummy variable $h$. After composing the raw coordinates with smooth
clipping maps whose images and first derivatives are globally bounded and
permutation-equivariant, the family \eqref{eq:psi-concrete} fits
\cref{ass:main}.
\par
Accordingly, the abstract results in the main text should be read as statements
about the presence or absence of \emph{boundary fast weights}. In the full
model \eqref{eq:model-full-matrix}--\eqref{eq:model-full}, the boundary state is
the pair $(s^{(1)},r^{(2)})$; in the regime without boundary fast weights there
is no boundary state. \Cref{prop:concrete-family} below records one convenient
set of sufficient conditions under which the clipped concrete family satisfies
\cref{ass:main}. \Cref{prop:identity-transfer} then records the precise
finite-horizon continuation principle needed to transfer those clipped results
back to the unclipped concrete model within the same bounded-input and
bounded-Lipschitz activation setting, once an a priori bound keeps the
trajectory inside an identity region of the clipping maps.
\end{remark}

\begin{proposition}[A clipped concrete BatchEnsemble family satisfies \cref{ass:main}]
\label{prop:concrete-family}
Assume the input domain is bounded:
\[
  \|x\|\le X_*
  \qquad
  \forall x\in\mathcal X,
\]
and assume that $\sigma$ and $\sigma'$ are globally bounded and globally
Lipschitz on $\R$. Let
\[
  T_\Theta:\Theta_{\mathrm{BE}}\to\Theta_{\mathrm{BE}},
  \qquad
  T_\mathcal H:\mathcal H_{\mathrm{BE}}\to\mathcal H_{\mathrm{BE}}
\]
be $C^1$ maps with globally bounded images, globally bounded and globally
Lipschitz first derivatives, and suppose they commute with head permutations.
Define
\[
  \widetilde\Psi_\alpha^{\mathrm{BE}}(x;\theta,h)
  :=
  \Psi_\alpha^{\mathrm{BE}}(x;T_\Theta(\theta),T_\mathcal H(h)).
\]
Then the feature family
\(
  (\widetilde\Psi_\alpha^{\mathrm{BE}})_{\alpha=1}^K
\)
satisfies \cref{ass:main}.
\end{proposition}

\begin{proposition}[Finite-horizon identity-region transfer to the raw concrete model]
\label{prop:identity-transfer}
Assume the bounded-input and activation hypotheses of
\cref{prop:concrete-family}. Suppose moreover that for each $R>0$ there are
maps
\[
  T_{\Theta,R}:\Theta_{\mathrm{BE}}\to\Theta_{\mathrm{BE}},
  \qquad
  T_{\mathcal H,R}:\mathcal H_{\mathrm{BE}}\to\mathcal H_{\mathrm{BE}}
\]
satisfying the assumptions of \cref{prop:concrete-family}, together with sets
\[
  \Theta_R\subseteq\Theta_{\mathrm{BE}},
  \qquad
  \mathcal H_R\subseteq\mathcal H_{\mathrm{BE}},
\]
such that for every $\theta\in\Theta_R$ and $h\in\mathcal H_R$,
\[
  T_{\Theta,R}(\theta)=\theta,
  \qquad
  \nabla T_{\Theta,R}(\theta)=I,
  \qquad
  T_{\mathcal H,R}(h)=h,
  \qquad
  \nabla T_{\mathcal H,R}(h)=I.
\]
Let $\widetilde\Psi_{\alpha,R}^{\mathrm{BE}}$ denote the clipped feature family
generated by $(T_{\Theta,R},T_{\mathcal H,R})$, and let
$(B_R,G_R)$ and $(B_{\mathrm{raw}},G_{\mathrm{raw}})$ be the corresponding
mean-field drifts for the clipped and raw concrete models, respectively.

If $\rho\in\PP_1(\Theta_{\mathrm{BE}})$ satisfies $\operatorname{supp}(\rho)\subseteq\Theta_R$
and $h\in\mathcal H_R$, then for every $\theta\in\Theta_R$,
\begin{align}
  B_R(\theta;\rho,h)
  &= B_{\mathrm{raw}}(\theta;\rho,h),
  \label{eq:transfer-B}\\
  G_R(\rho,h)
  &= G_{\mathrm{raw}}(\rho,h).
  \label{eq:transfer-G}
\end{align}
Consequently, if either the clipped or the raw concrete characteristic
solution with some common initial data remains inside
\[
  \bigl\{
    (\rho,h)\in\PP_1(\Theta_{\mathrm{BE}})\times\mathcal H_{\mathrm{BE}}
    :
    \operatorname{supp}(\rho)\subseteq\Theta_R,\ h\in\mathcal H_R
  \bigr\}
\]
throughout $[0,T]$, then on $[0,T]$ it is simultaneously a solution of the
other system as well.
\end{proposition}

\begin{remark}[What determines the transferable horizon?]
\label{rem:transfer-horizon}
For a chosen family of identity regions $(\Theta_R,\mathcal H_R)$, the
relevant quantity is the first exit time
\[
  \tau_R
  :=
  \inf\bigl\{
    t\ge 0
    :
    \operatorname{supp}(\rho_t)\nsubseteq\Theta_R
    \text{ or }
    h_t\notin\mathcal H_R
  \bigr\}.
\]
\Cref{prop:identity-transfer} upgrades the clipped theory to the raw concrete
BatchEnsemble model on every interval $[0,T]$ with $T<\tau_R$, still within the
same bounded-input and globally bounded/Lipschitz activation setting used to
construct the clipped theory. Extending this step to unbounded or nonsmooth
positively homogeneous activations, such as ReLU, would require additional
arguments and is not claimed here. Thus one does not need a global-in-time
boundedness theorem to remove clipping on practical training horizons inside
that hypothesis class; it is enough to prove that, for each prescribed finite
$T>0$, there exists some radius $R(T)$ such that $\tau_{R(T)}>T$.

A natural route is an a priori finite-horizon moment estimate for the raw
system. For example, one may aim to control quantities of the form
\[
  \sup_{t\in[0,T]}
  \Biggl[
    \int_{\Theta_{\mathrm{BE}}}
    \Bigl(
      \|a\|^2+\|w\|^2
      +\sum_{\alpha=1}^K
      \bigl(
        |s_\alpha^{(2)}|^2+|r_\alpha^{(1)}|^2
      \bigr)
    \Bigr)\rho_t(\dd\theta)
    +
    \sum_{\alpha=1}^K
    \bigl(
      \|u_{\alpha,t}\|^2+\|v_{\alpha,t}\|^2
    \bigr)
  \Biggr].
\]
Such a bound would be especially plausible under bounded-support
initializations for the boundary variables, such as random-sign or compactly
supported uniform laws. It is a finite-horizon continuation statement, not a
claim of global-in-time boundedness: because of scale-separation and
margin-seeking directions, one should generally not expect the raw parameter
trajectory to remain uniformly bounded as $t\to\infty$ even when the predictor
itself stays well controlled.
\end{remark}

\subsection{General Fixed-\texorpdfstring{$K$}{K} Width Limit}
\label{sec:appendix-general-limit}

\begin{theorem}[Deterministic well-posedness, stability, and particle convergence]
\label{thm:deterministic}
Assume \cref{ass:main}, fix $T>0$, and let $(\rho_0,h_0)\in\PP_1(\Theta)\times\mathcal H$.

\begin{enumerate}[label=(\roman*),leftmargin=2em]
  \item There exists a unique characteristic solution
  \[
    (\rho_t,h_t)_{t\in[0,T]}
    \in
    C\bigl([0,T],\PP_1(\Theta)\times\mathcal H\bigr)
  \]
  of \cref{eq:pde,eq:ode}.
  \item There exists a constant $C_T<\infty$ such that for any two initial data
  $(\rho_0,h_0)$ and $(\mu_0,\bar h_0)$ with corresponding solutions
  $(\rho_t,h_t)$ and $(\mu_t,\bar h_t)$,
  \begin{equation}
    \sup_{t\in[0,T]}
    \bigl(
      \W(\rho_t,\mu_t)+\|h_t-\bar h_t\|
    \bigr)
    \le
    C_T
    \bigl(
      \W(\rho_0,\mu_0)+\|h_0-\bar h_0\|
    \bigr).
    \label{eq:det-stability}
  \end{equation}
  \item Let deterministic particles and boundary states satisfy
  \begin{align*}
    \dot\theta_{m,t}^M &= B(\theta_{m,t}^M;\rho_t^M,h_t^M),
    \qquad
    \rho_t^M=\frac1M\sum_{m=1}^M\delta_{\theta_{m,t}^M},\\
    \dot h_t^M &= G(\rho_t^M,h_t^M).
  \end{align*}
  If
  \[
    \W(\rho_0^M,\rho_0)+\|h_0^M-h_0\|\xrightarrow[M\to\infty]{}0,
  \]
  then
  \begin{equation}
    \sup_{t\in[0,T]}
    \bigl(
      \W(\rho_t^M,\rho_t)+\|h_t^M-h_t\|
    \bigr)
    \le
    C_T\bigl(\W(\rho_0^M,\rho_0)+\|h_0^M-h_0\|\bigr)
    \xrightarrow[M\to\infty]{}0.
    \label{eq:particle-conv}
  \end{equation}
  In particular,
  \[
    \sup_{t\in[0,T]}
    \|f_\alpha(x;\rho_t^M,h_t^M)-f_\alpha(x;\rho_t,h_t)\|
    \to 0
  \]
  for every fixed $\alpha$ and $x$.
\end{enumerate}
\end{theorem}

\begin{corollary}[Random initialization]
\label{cor:random}
Assume \cref{ass:main}, fix $T>0$, and let $(\rho_0,h_0)$ be a possibly random
initial environment. Suppose the random initial data $(\rho_0^M,h_0^M)$ satisfy
\[
  \W(\rho_0^M,\rho_0)+\|h_0^M-h_0\|\xrightarrow[M\to\infty]{}0
\]
in probability, almost surely, or in $L^1$. Let $(\rho_t^M,h_t^M)$ be the
finite-width particle/boundary system and let $(\rho_t,h_t)$ be the characteristic
solution of \cref{eq:pde,eq:ode} started from $(\rho_0,h_0)$. Then the same mode
of convergence holds uniformly on $[0,T]$:
\[
  \sup_{t\in[0,T]}
  \bigl(
    \W(\rho_t^M,\rho_t)+\|h_t^M-h_t\|
  \bigr)
  \to 0.
\]
Consequently, the infinite-width predictor vector is a measurable function of the
initial environment $(\rho_0,h_0)$.
\end{corollary}

\section{Proofs}
\label{sec:appendix-proofs}

\subsection{Clipped Concrete Family}

\begin{proof}
Write
\[
  T_\Theta(\theta)
  =
  (\widetilde a,\widetilde w,\widetilde s_1^{(2)},\dots,\widetilde s_K^{(2)},
  \widetilde r_1^{(1)},\dots,\widetilde r_K^{(1)})
\]
and
\[
  T_\mathcal H(h)=(\widetilde u,\widetilde v),
  \qquad
  \widetilde u=(\widetilde u_1,\dots,\widetilde u_K),
  \qquad
  \widetilde v=(\widetilde v_1,\dots,\widetilde v_K).
\]
Then
\[
  \widetilde\Psi_\alpha^{\mathrm{BE}}(x;\theta,h)
  =
  (\widetilde a\odot \widetilde v_\alpha)\,\widetilde s_\alpha^{(2)}\,
  \sigma\!\bigl(
    \widetilde r_\alpha^{(1)}\,\widetilde w^\top(\widetilde u_\alpha\odot x)
  \bigr).
\]
Because $T_\Theta$ and $T_\mathcal H$ have bounded images, every clipped
coordinate above is uniformly bounded. Since $\|x\|\le X_*$, the pre-activation
\[
  z_\alpha(\theta,h;x)
  :=
  \widetilde r_\alpha^{(1)}\,\widetilde w^\top(\widetilde u_\alpha\odot x)
\]
is uniformly bounded in $(\theta,h,x)$. The clipped coordinates are globally
Lipschitz, so $z_\alpha(\cdot,\cdot;x)$ is globally Lipschitz in $(\theta,h)$
uniformly in $x$.

By assumption, $\sigma$ and $\sigma'$ are globally bounded and globally
Lipschitz. Hence both $\sigma(z_\alpha)$ and $\sigma'(z_\alpha)$ are uniformly
bounded and globally Lipschitz in $(\theta,h)$, uniformly in $x$. Since
$\widetilde\Psi_\alpha^{\mathrm{BE}}$ is a product of bounded Lipschitz factors,
it is itself uniformly bounded and globally Lipschitz in $(\theta,h)$.

Next, differentiate with respect to $\theta$ and $h_\alpha$. By the chain rule,
every component of $\nabla_\theta\widetilde\Psi_\alpha^{\mathrm{BE}}$ and
$\nabla_{h_\alpha}\widetilde\Psi_\alpha^{\mathrm{BE}}$ is a finite sum of terms
obtained by multiplying:
\begin{enumerate}[label=(\roman*),leftmargin=2em]
  \item bounded clipped coordinates,
  \item bounded first derivatives of $T_\Theta$ or $T_\mathcal H$,
  \item one copy of $\sigma(z_\alpha)$ or $\sigma'(z_\alpha)$,
  \item and, in the $\sigma'(z_\alpha)$ terms, bounded linear forms in $x$ such
  as $\widetilde w^\top(\widetilde u_\alpha\odot x)$ or coordinates of
  $\widetilde u_\alpha\odot x$.
\end{enumerate}
All these factors are uniformly bounded and globally Lipschitz in $(\theta,h)$,
again uniformly in $x$, so the same is true of the first derivatives. This
verifies the bounded-Lipschitz part of \cref{ass:main}.

For permutation equivariance, let $\pi\in S_K$. The unclipped concrete feature
family already satisfies
\[
  \Psi_\alpha^{\mathrm{BE}}(x;\Pi_\pi^\Theta\theta,\Pi_\pi^\mathcal H h)
  =
  \Psi_{\pi^{-1}(\alpha)}^{\mathrm{BE}}(x;\theta,h),
\]
because $\Pi_\pi^\Theta$ and $\Pi_\pi^\mathcal H$ simply permute the
head-labelled coordinates. Since $T_\Theta$ and $T_\mathcal H$ commute with the
same permutations, the clipped family satisfies the same identity. Therefore
\cref{ass:main} holds for
\(
  (\widetilde\Psi_\alpha^{\mathrm{BE}})_{\alpha=1}^K
\).
\end{proof}

\begin{proof}[Proof of \cref{prop:identity-transfer}]
Fix $R>0$, let $\rho\in\PP_1(\Theta_{\mathrm{BE}})$ satisfy
$\operatorname{supp}(\rho)\subseteq\Theta_R$, and let $h\in\mathcal H_R$. For every
$\theta\in\Theta_R$, the identities
\[
  T_{\Theta,R}(\theta)=\theta,
  \qquad
  \nabla T_{\Theta,R}(\theta)=I,
  \qquad
  T_{\mathcal H,R}(h)=h,
  \qquad
  \nabla T_{\mathcal H,R}(h)=I
\]
imply, by the chain rule, that for every head $\alpha$ and input $x$,
\[
\begin{aligned}
  \widetilde\Psi_{\alpha,R}^{\mathrm{BE}}(x;\theta,h)
  &=
  \Psi_\alpha^{\mathrm{BE}}(x;\theta,h),\\
  \nabla_\theta\widetilde\Psi_{\alpha,R}^{\mathrm{BE}}(x;\theta,h)
  &=
  \nabla_\theta\Psi_\alpha^{\mathrm{BE}}(x;\theta,h),\\
  \nabla_{h_\alpha}\widetilde\Psi_{\alpha,R}^{\mathrm{BE}}(x;\theta,h)
  &=
  \nabla_{h_\alpha}\Psi_\alpha^{\mathrm{BE}}(x;\theta,h).
\end{aligned}
\]
Integrating the feature identity against $\rho$ yields equality of the clipped
and raw outputs:
\[
  \widetilde f_{\alpha,R}(x;\rho,h)
  =
  f_\alpha^{\mathrm{raw}}(x;\rho,h)
  \qquad
  \forall \alpha,\ \forall x\in\mathcal X.
\]
Substituting these identities into the drift definitions shows that the clipped
and raw drifts coincide, which is exactly \eqref{eq:transfer-B} and
\eqref{eq:transfer-G}.

For the last claim, consider either the raw or clipped characteristic solution
with the given initial data and assume it remains in the identity region on
$[0,T]$. Along that trajectory the two drift fields agree at every time, so the
same path satisfies the other evolution system as well.
\end{proof}

\subsection{Auxiliary Lipschitz Bounds}

\begin{proposition}[Output and drift Lipschitz bounds]
\label{prop:lipschitz}
Assume \cref{ass:main}. Then for every $x\in\mathcal X$ and every $\alpha$,
\begin{equation}
  \|f_\alpha(x;\rho,h)-f_\alpha(x;\mu,\bar h)\|
  \le
  L_\Psi\bigl(\W(\rho,\mu)+\|h-\bar h\|\bigr).
  \label{eq:out-lip}
\end{equation}
Moreover, there exist finite constants $L_B,L_G$ depending only on $K$ and on
the constants in \cref{ass:main} such that
\begin{align}
  \|B(\theta;\rho,h)-B(\eta;\mu,\bar h)\|
  &\le
  L_B\bigl(
    \|\theta-\eta\|+\W(\rho,\mu)+\|h-\bar h\|
  \bigr),
  \label{eq:B-lip}\\
  \|G(\rho,h)-G(\mu,\bar h)\|
  &\le
  L_G\bigl(\W(\rho,\mu)+\|h-\bar h\|\bigr).
  \label{eq:G-lip}
\end{align}
\end{proposition}

\begin{proof}
For the output bound, fix any coupling $\gamma$ of $\rho$ and $\mu$. Then
\begin{align*}
  \|f_\alpha(x;\rho,h)-f_\alpha(x;\mu,\bar h)\|
  &\le
  \int_{\Theta\times\Theta}
  \|\Psi_\alpha(x;\theta,h)-\Psi_\alpha(x;\eta,\bar h)\|
  \,\gamma(\dd\theta,\dd\eta)\\
  &\le
  L_\Psi
  \int_{\Theta\times\Theta}
  \bigl(\|\theta-\eta\|+\|h-\bar h\|\bigr)\,\gamma(\dd\theta,\dd\eta).
\end{align*}
Taking the infimum over $\gamma$ gives \eqref{eq:out-lip}.

For the neuron drift, write
\[
  B(\theta;\rho,h)-B(\eta;\mu,\bar h)=I_1+I_2,
\]
where
\begin{align*}
  I_1
  &:=
  -\E\Bigl[
    \sum_{\alpha=1}^K
    \nabla_\theta\Psi_\alpha(x;\theta,h)^\top
    \bigl(
      \nabla_1\ell(f_\alpha(x;\rho,h),y)
      -
      \nabla_1\ell(f_\alpha(x;\mu,\bar h),y)
    \bigr)
  \Bigr],\\
  I_2
  &:=
  -\E\Bigl[
    \sum_{\alpha=1}^K
    \bigl(
      \nabla_\theta\Psi_\alpha(x;\theta,h)
      -
      \nabla_\theta\Psi_\alpha(x;\eta,\bar h)
    \bigr)^\top
    \nabla_1\ell(f_\alpha(x;\mu,\bar h),y)
  \Bigr].
\end{align*}
Using \eqref{eq:out-lip}, the Lipschitz property of $\nabla_1\ell$, and the
uniform bound on $\nabla_\theta\Psi_\alpha$ gives
\[
  \|I_1\|
  \le
  K L_\ell B_{\nabla\Psi}L_\Psi
  \bigl(\W(\rho,\mu)+\|h-\bar h\|\bigr).
\]
Because \cref{ass:main}(i) implies
\(
  \|f_\alpha(x;\mu,\bar h)\|\le B_\Psi
\)
for every head, the local envelope bound \eqref{eq:loss-envelope} and the
Lipschitz property of $\nabla_\theta\Psi_\alpha$ give
\[
  \|I_2\|
  \le
  K C_{\ell,\Psi} L_{\nabla\Psi}
  \bigl(\|\theta-\eta\|+\|h-\bar h\|\bigr).
\]
Therefore \eqref{eq:B-lip} holds with
\[
  L_B:=K\bigl(L_\ell B_{\nabla\Psi}L_\Psi+C_{\ell,\Psi} L_{\nabla\Psi}\bigr).
\]

The proof of \eqref{eq:G-lip} is the same. One writes
$G(\rho,h)-G(\mu,\bar h)=J_1+J_2$, where $J_1$ contains the difference of the
loss gradients and $J_2$ the difference of
$\nabla_{h_\alpha}\Psi_\alpha(\cdot,\cdot)$, and then uses \eqref{eq:out-lip},
the bound on $\nabla_{h_\alpha}\Psi_\alpha$, the local loss envelope
\eqref{eq:loss-envelope}, and the Lipschitz property of
$\nabla_{h_\alpha}\Psi_\alpha$. The result is \eqref{eq:G-lip} with
\[
  L_G:=K\bigl(L_\ell B_{\nabla\Psi}L_\Psi+C_{\ell,\Psi} L_{\nabla\Psi}\bigr).
\]
\end{proof}

\subsection{Deterministic Width Limit}

\begin{proof}[Proof of \cref{thm:deterministic}]
Fix $T>0$ and endow
\[
  \mathcal C_T
  :=
  C\bigl([0,T],\PP_1(\Theta)\times\mathcal H\bigr)
\]
with the metric
\[
  d_T\bigl((\rho,h),(\mu,\bar h)\bigr)
  :=
  \sup_{t\in[0,T]}
  \bigl(
    \W(\rho_t,\mu_t)+\|h_t-\bar h_t\|
  \bigr).
\]

\paragraph{Step 1: existence and uniqueness on a short interval.}
Given $(\widehat\rho,\widehat h)\in\mathcal C_T$, let $\Phi_t^{\widehat\rho,\widehat h}$
solve
\[
  \dot\Phi_t^{\widehat\rho,\widehat h}(\theta)
  =
  B(\Phi_t^{\widehat\rho,\widehat h}(\theta);\widehat\rho_t,\widehat h_t),
  \qquad
  \Phi_0^{\widehat\rho,\widehat h}(\theta)=\theta.
\]
By \cref{prop:lipschitz}, the map
$\theta\mapsto B(\theta;\widehat\rho_t,\widehat h_t)$ is globally Lipschitz
uniformly in $t$, so the flow exists uniquely for all $t\in[0,T]$. Moreover,
\cref{ass:main}(i) gives
\(
  \|f_\alpha(x;\rho,h)\|\le B_\Psi
\)
for every head, so by \eqref{eq:loss-envelope},
\[
  \|B(\theta;\rho,h)\|
  \le
  \E\Bigl[
    \sum_{\alpha=1}^K
    \|\nabla_\theta\Psi_\alpha(x;\theta,h)\|\,
    \|\nabla_1\ell(f_\alpha(x;\rho,h),y)\|
  \Bigr]
  \le
  K B_{\nabla\Psi}C_{\ell,\Psi},
\]
so
\[
  \|\Phi_t^{\widehat\rho,\widehat h}(\theta)\|
  \le
  \|\theta\|+K B_{\nabla\Psi}C_{\ell,\Psi} t.
\]
Hence $(\Phi_t^{\widehat\rho,\widehat h})_\#\rho_0$ has finite first moment for
all $t\in[0,T]$.

Define the operator $\mathcal T:\mathcal C_T\to\mathcal C_T$ by
\[
  \mathcal T(\widehat\rho,\widehat h)_t
  :=
  \Bigl(
    (\Phi_t^{\widehat\rho,\widehat h})_\#\rho_0,\,
    h_0+\int_0^t G(\widehat\rho_s,\widehat h_s)\,\dd s
  \Bigr).
\]
Take two curves $(\widehat\rho,\widehat h)$ and $(\widehat\mu,\widehat{\bar h})$
in $\mathcal C_T$, and write $\Phi_t$ and $\Psi_t$ for the corresponding flows.
For each $\theta\in\Theta$,
\[
  \frac{\dd}{\dd t}\|\Phi_t(\theta)-\Psi_t(\theta)\|
  \le
  L_B\|\Phi_t(\theta)-\Psi_t(\theta)\|
  +
  L_B
  \bigl(
    \W(\widehat\rho_t,\widehat\mu_t)
    +
    \|\widehat h_t-\widehat{\bar h}_t\|
  \bigr).
\]
Gronwall's inequality gives
\[
  \|\Phi_t(\theta)-\Psi_t(\theta)\|
  \le
  \bigl(e^{L_B t}-1\bigr)\,
  d_T\bigl((\widehat\rho,\widehat h),(\widehat\mu,\widehat{\bar h})\bigr).
\]
Using the coupling induced by the common initial law $\rho_0$, we obtain
\[
  \W\bigl(
    (\Phi_t)_\#\rho_0,
    (\Psi_t)_\#\rho_0
  \bigr)
  \le
  \bigl(e^{L_B t}-1\bigr)\,
  d_T\bigl((\widehat\rho,\widehat h),(\widehat\mu,\widehat{\bar h})\bigr).
\]
For the boundary component,
\begin{align*}
  \Bigl\|
    \int_0^t G(\widehat\rho_s,\widehat h_s)\,\dd s
    -
    \int_0^t G(\widehat\mu_s,\widehat{\bar h}_s)\,\dd s
  \Bigr\|
  &\le
  \int_0^t
  \|G(\widehat\rho_s,\widehat h_s)-G(\widehat\mu_s,\widehat{\bar h}_s)\|
  \,\dd s\\
  &\le
  T L_G\,
  d_T\bigl((\widehat\rho,\widehat h),(\widehat\mu,\widehat{\bar h})\bigr).
\end{align*}
Therefore
\[
  d_T\bigl(
    \mathcal T(\widehat\rho,\widehat h),
    \mathcal T(\widehat\mu,\widehat{\bar h})
  \bigr)
  \le
  q_T\,
  d_T\bigl((\widehat\rho,\widehat h),(\widehat\mu,\widehat{\bar h})\bigr),
\]
where
\[
  q_T:=e^{L_B T}-1+T L_G.
\]
For $T>0$ small enough, $q_T<1$, so $\mathcal T$ is a contraction. Banach's
fixed-point theorem yields a unique characteristic solution on such a short
interval. Iterating this argument over finitely many subintervals gives a
unique characteristic solution on any prescribed finite horizon.

\paragraph{Step 2: stability.}
Let $(\rho_t,h_t)$ and $(\mu_t,\bar h_t)$ be the characteristic solutions
started from $(\rho_0,h_0)$ and $(\mu_0,\bar h_0)$, respectively, and let
$\Phi_t$ and $\bar\Phi_t$ be the corresponding characteristic flows. Fix any
coupling $\gamma_0$ of $\rho_0$ and $\mu_0$, and set
\[
  \gamma_t:=(\Phi_t,\bar\Phi_t)_\#\gamma_0.
\]
Then $\gamma_t$ is a coupling of $\rho_t$ and $\mu_t$. Define
\[
  D_t^\Theta
  :=
  \int_{\Theta\times\Theta}
  \|\Phi_t(\theta)-\bar\Phi_t(\eta)\|
  \,\gamma_0(\dd\theta,\dd\eta),
  \qquad
  D_t^h:=\|h_t-\bar h_t\|.
\]
Since $\W(\rho_t,\mu_t)\le D_t^\Theta$, \eqref{eq:B-lip} implies
\begin{align*}
  \frac{\dd}{\dd t}D_t^\Theta
  &\le
  \int_{\Theta\times\Theta}
  \|B(\Phi_t(\theta);\rho_t,h_t)-B(\bar\Phi_t(\eta);\mu_t,\bar h_t)\|
  \,\gamma_0(\dd\theta,\dd\eta)\\
  &\le
  L_B\bigl(D_t^\Theta+\W(\rho_t,\mu_t)+D_t^h\bigr)\\
  &\le
  2L_B D_t^\Theta + L_B D_t^h.
\end{align*}
Similarly, \eqref{eq:G-lip} gives
\[
  \frac{\dd}{\dd t}D_t^h
  \le
  L_G\bigl(\W(\rho_t,\mu_t)+D_t^h\bigr)
  \le
  L_G(D_t^\Theta+D_t^h).
\]
With $D_t:=D_t^\Theta+D_t^h$, we obtain
\[
  \dot D_t
  \le
  (2L_B+L_G)D_t.
\]
Hence
\[
  D_t
  \le
  e^{(2L_B+L_G)t}D_0
  =
  e^{(2L_B+L_G)t}
  \Bigl(
    \int_{\Theta\times\Theta}\|\theta-\eta\|\,\gamma_0(\dd\theta,\dd\eta)
    +
    \|h_0-\bar h_0\|
  \Bigr).
\]
Taking the infimum over all couplings $\gamma_0$ proves \eqref{eq:det-stability}
with
\[
  C_T:=e^{(2L_B+L_G)T}.
\]

\paragraph{Step 3: particle convergence.}
The particle system generates a flow $\Phi_t^M$ satisfying
\[
  \dot\Phi_t^M(\theta_{m,0}^M)
  =
  B(\Phi_t^M(\theta_{m,0}^M);\rho_t^M,h_t^M),
  \qquad
  \rho_t^M=(\Phi_t^M)_\#\rho_0^M,
\]
while $h_t^M$ solves
\[
  \dot h_t^M=G(\rho_t^M,h_t^M).
\]
Thus $(\rho_t^M,h_t^M)$ is itself the characteristic solution of
\cref{eq:pde,eq:ode} with initial data $(\rho_0^M,h_0^M)$. Applying the
stability estimate from Step 2 with
\(
  (\mu_t,\bar h_t)=(\rho_t^M,h_t^M)
\)
gives \eqref{eq:particle-conv}. The predictor convergence then follows from
\eqref{eq:out-lip}.
\end{proof}

\subsection{Random Initialization}

\begin{proof}[Proof of \cref{cor:random}]
Fix $\omega$ in the underlying sample space. Conditional on that $\omega$, the
initial data $(\rho_0^M(\omega),h_0^M(\omega))$ and $(\rho_0(\omega),h_0(\omega))$
are deterministic, so \cref{thm:deterministic} yields
\[
  \sup_{t\in[0,T]}
  \bigl(
    \W(\rho_t^M,\rho_t)+\|h_t^M-h_t\|
  \bigr)
  \le
  C_T
  \bigl(
    \W(\rho_0^M,\rho_0)+\|h_0^M-h_0\|
  \bigr)
\]
pointwise in $\omega$.
Therefore convergence in probability and almost sure convergence are inherited
directly from the corresponding convergence at time $0$. If the initial
convergence holds in $L^1$, then taking expectations gives
\[
  \E\Bigl[
    \sup_{t\in[0,T]}
    \bigl(
      \W(\rho_t^M,\rho_t)+\|h_t^M-h_t\|
    \bigr)
  \Bigr]
  \le
  C_T
  \E\bigl[
    \W(\rho_0^M,\rho_0)+\|h_0^M-h_0\|
  \bigr]
  \xrightarrow[M\to\infty]{}0.
\]

The same estimate also shows that the solution map
\[
  S_t:(\rho_0,h_0)\mapsto(\rho_t,h_t)
\]
is continuous from $\PP_1(\Theta)\times\mathcal H$ to itself for every fixed
$t$. Since $(\rho_t,h_t)=S_t(\rho_0,h_0)$ and
\(
  f_\alpha(x;\rho_t,h_t)
\)
is continuous in $(\rho_t,h_t)$ by \eqref{eq:out-lip}, the infinite-width
predictor vector is a measurable function of the initial environment
$(\rho_0,h_0)$.
\end{proof}

\subsection{Symmetric-Initialization Collapse}

\begin{proof}[Proof of \cref{thm:identical-boundary-collapse}]
By \cref{thm:deterministic}, it is enough to prove that the permutation-transformed
pair is another characteristic solution with the same initial data.

Fix $\pi\in S_K$, and write
\[
  Q:=\Pi_\pi^\Theta,
  \qquad
  P:=\Pi_\pi^\mathcal H.
\]
Define
\[
  \widetilde\rho_t:=Q_\#\rho_t,
  \qquad
  \widetilde h_t:=P h_t.
\]
We first record the permutation identities for outputs and drifts.

For the outputs, a change of variables together with \eqref{eq:equiv} gives
\begin{align}
  f_\alpha(x;\widetilde\rho_t,\widetilde h_t)
  &=
  \int_\Theta \Psi_\alpha(x;\theta,\widetilde h_t)\,\widetilde\rho_t(\dd\theta)
  \notag\\
  &=
  \int_\Theta \Psi_\alpha(x;Q\eta,P h_t)\,\rho_t(\dd\eta)
  \notag\\
  &=
  \int_\Theta \Psi_{\pi^{-1}(\alpha)}(x;\eta,h_t)\,\rho_t(\dd\eta)
  =
  f_{\pi^{-1}(\alpha)}(x;\rho_t,h_t).
  \label{eq:perm-output}
\end{align}

For the neuron drift, differentiate \eqref{eq:equiv} in $\theta$.
Since $Q$ is a linear isometry, $Q^{-1}=Q^\top$, and therefore
\[
  \nabla_\theta\Psi_\alpha(x;Q\theta,P h)^\top
  =
  Q\,
  \nabla_\theta\Psi_{\pi^{-1}(\alpha)}(x;\theta,h)^\top.
\]
Combining this with \eqref{eq:perm-output} and reindexing the head sum gives
\begin{align}
  B(Q\theta;\widetilde\rho_t,\widetilde h_t)
  &=
  -\E\Bigl[
    \sum_{\alpha=1}^K
    \nabla_\theta\Psi_\alpha(x;Q\theta,P h_t)^\top
    \nabla_1\ell(f_\alpha(x;\widetilde\rho_t,\widetilde h_t),y)
  \Bigr]
  \notag\\
  &=
  -Q\E\Bigl[
    \sum_{\alpha=1}^K
    \nabla_\theta\Psi_{\pi^{-1}(\alpha)}(x;\theta,h_t)^\top
    \nabla_1\ell(f_{\pi^{-1}(\alpha)}(x;\rho_t,h_t),y)
  \Bigr]
  \notag\\
  &=
  Q\,B(\theta;\rho_t,h_t).
  \label{eq:perm-B}
\end{align}

For the boundary drift, let $e_\alpha(u)\in\mathcal H$ denote the vector whose
$\alpha$-th block is $u\in\R^{d_h}$ and whose other blocks vanish. Since
$P e_{\pi^{-1}(\alpha)}(u)=e_\alpha(u)$, \eqref{eq:equiv} implies
\[
  \Psi_\alpha(x;Q\theta,P h+t e_\alpha(u))
  =
  \Psi_{\pi^{-1}(\alpha)}(x;\theta,h+t e_{\pi^{-1}(\alpha)}(u)).
\]
Differentiating at $t=0$ yields
\[
  \nabla_{h_\alpha}\Psi_\alpha(x;Q\theta,P h)^\top u
  =
  \nabla_{h_{\pi^{-1}(\alpha)}}
  \Psi_{\pi^{-1}(\alpha)}(x;\theta,h)^\top u,
  \qquad
  \forall u\in\R^{d_h}.
\]
Hence
\[
  \nabla_{h_\alpha}\Psi_\alpha(x;Q\theta,P h)^\top
  =
  \nabla_{h_{\pi^{-1}(\alpha)}}
  \Psi_{\pi^{-1}(\alpha)}(x;\theta,h)^\top,
\]
and therefore
\[
  G_\alpha(\widetilde\rho_t,\widetilde h_t)
  =
  G_{\pi^{-1}(\alpha)}(\rho_t,h_t).
\]
Equivalently,
\begin{equation}
  G(\widetilde\rho_t,\widetilde h_t)=P\,G(\rho_t,h_t).
  \label{eq:perm-G}
\end{equation}

Now let $\Phi_t$ be the characteristic flow of $(\rho_t,h_t)$ and define
\[
  \widetilde\Phi_t:=Q\circ \Phi_t\circ Q^{-1}.
\]
Using \eqref{eq:perm-B},
\[
  \frac{\dd}{\dd t}\widetilde\Phi_t(\theta)
  =
  Q\,B(Q^{-1}\widetilde\Phi_t(\theta);\rho_t,h_t)
  =
  B(\widetilde\Phi_t(\theta);\widetilde\rho_t,\widetilde h_t).
\]
Also,
\[
  \widetilde\rho_t=(\widetilde\Phi_t)_\#\widetilde\rho_0,
  \qquad
  \dot{\widetilde h}_t
  =
  P\dot h_t
  =
  P\,G(\rho_t,h_t)
  =
  G(\widetilde\rho_t,\widetilde h_t)
\]
by \eqref{eq:perm-G}. Thus $(\widetilde\rho_t,\widetilde h_t)$ is the
characteristic solution of \cref{eq:pde,eq:ode} started from
\[
  (\widetilde\rho_0,\widetilde h_0)
  =
  (Q_\#\rho_0,P h_0).
\]
If the initial data are jointly permutation-invariant, then
$(\widetilde\rho_0,\widetilde h_0)=(\rho_0,h_0)$, so uniqueness from
\cref{thm:deterministic} yields
\[
  Q_\#\rho_t=\rho_t,
  \qquad
  P h_t=h_t
  \qquad
  \forall t\ge 0.
\]

Finally, fix $\alpha,\beta\in\{1,\dots,K\}$ and choose $\pi$ such that
$\pi^{-1}(\alpha)=\beta$. Using \eqref{eq:perm-output} and the just-proved
invariance,
\[
  f_\alpha(x;\rho_t,h_t)
  =
  f_\alpha(x;Q_\#\rho_t,P h_t)
  =
  f_{\pi^{-1}(\alpha)}(x;\rho_t,h_t)
  =
  f_\beta(x;\rho_t,h_t).
\]
This proves head collapse under identical boundary initialization.
\end{proof}

\subsection{Boundary-State Randomness}

\begin{proof}[Proof of \cref{cor:boundary-init}]
If $(\rho_0,h_0)$ is deterministic, then \cref{cor:random} shows that the
entire limit trajectory $(\rho_t,h_t)_{t\in[0,T]}$ is deterministic, hence so is
the predictor vector. If $\rho_0$ is deterministic but $h_0$ is random, then
\cref{cor:random} gives
\[
  (\rho_t,h_t)=S_t(\rho_0,h_0)
\]
for a measurable solution map $S_t$. Since the first argument is fixed, the
limit predictor vector is a measurable function of $h_0$ alone. Thus any
order-one randomness surviving in the infinite-width limit is induced entirely
by the law of $h_0$.
\end{proof}

\subsection{Boundary-Only Common Flow}

\begin{proof}
For the boundary-only model \eqref{eq:model-bdry}, write
\[
  q=(u,v)\in\R^{d_x}\times\R^{d_y},
  \qquad
  \widehat\Psi(x;(a,w),q)
  :=
  (a\odot v)\,\sigma\!\bigl(w^\top(u\odot x)\bigr).
\]
Then the feature family has the special form
\[
  \Psi_\alpha(x;\theta,h)=\widehat\Psi(x;\theta,h_\alpha),
  \qquad
  \theta=(a,w),
  \qquad
  h=(h_1,\dots,h_K).
\]
Consequently,
\[
  f_\alpha(x;\rho,h)
  =
  \int \widehat\Psi(x;\theta,h_\alpha)\,\rho(\dd\theta)
  =
  H(x;\rho,h_\alpha).
\]
Substituting this representation into \eqref{eq:drift-G} gives
\[
  G_\alpha(\rho,h)
  =
  -\E_{(x,y)\sim\mathsf P}
  \Bigl[
    \int
    \nabla_q\widehat\Psi(x;\theta,h_\alpha)^\top
    \nabla_1\ell(H(x;\rho,h_\alpha),y)\,\rho(\dd\theta)
  \Bigr],
\]
which is exactly \eqref{eq:boundary-drift} evaluated at
$q=h_\alpha$. Hence
\[
  \dot h_{\alpha,t}=G_\alpha(\rho_t,h_t)=C(\rho_t,h_{\alpha,t}),
  \qquad
  \alpha=1,\dots,K.
\]

Now let $\Phi_{s,t}$ be the two-parameter flow generated by the
nonautonomous ODE
\[
  \dot q_r=C(\rho_r,q_r),
  \qquad
  r\in[s,t].
\]
By uniqueness,
\[
  h_{\alpha,t}=\Phi_{0,t}(h_{\alpha,0}).
\]
If $h_{\alpha,0}=h_{\beta,0}$, then uniqueness immediately yields
$h_{\alpha,t}=h_{\beta,t}$ for every $t$.

For arbitrary initial states, the Lipschitz condition implies by Gronwall that
\[
  \|\Phi_{s,t}(q)-\Phi_{s,t}(\bar q)\|
  \le
  \exp\!\Bigl(\int_s^t L_r\,\dd r\Bigr)\|q-\bar q\|
\]
for all $q,\bar q$. Since $\Phi_{t,s}=(\Phi_{s,t})^{-1}$ is the flow of the
same vector field run backward on the finite interval $[s,t]$, the same bound
applies to the inverse map:
\[
  \|q-\bar q\|
  =
  \|\Phi_{t,s}(\Phi_{s,t}(q))-\Phi_{t,s}(\Phi_{s,t}(\bar q))\|
  \le
  \exp\!\Bigl(\int_s^t L_r\,\dd r\Bigr)
  \|\Phi_{s,t}(q)-\Phi_{s,t}(\bar q)\|.
 \]
Combining the two estimates and setting $s=0$,
$q=h_{\alpha,0}$, and $\bar q=h_{\beta,0}$ yields
\eqref{eq:boundary-bilipschitz}. The final non-coalescence statement follows
immediately.
\end{proof}

\subsection{Collapsed Boundary-Only Single-Model Flow}

\begin{proof}[Proof of \cref{prop:boundary-single-flow}]
By \cref{prop:boundary-flow}, identical boundary initialization stays identical,
so there exists a path
\[
  q_t=(u_t,v_t)\in\R^{d_x}\times\R^{d_y}
\]
such that
\[
  h_{1,t}=\cdots=h_{K,t}=q_t
  \qquad
  \forall t\ge 0.
\]
For the boundary-only model write
\[
  \widehat\Psi(x;(a,w),q)
  :=
  (a\odot v)\,\sigma\!\bigl(w^\top(u\odot x)\bigr).
\]
Then
\[
  \Psi_\alpha(x;\theta,h)=\widehat\Psi(x;\theta,h_\alpha),
  \qquad
  f_\alpha(x;\rho_t,h_t)=H(x;\rho_t,q_t)
\]
for every head $\alpha$. Substituting this into the neuron drift
\eqref{eq:drift-B} gives
\begin{align*}
  B((a,w);\rho_t,h_t)
  &=
  -\E_{(x,y)\sim\mathsf P}
  \Bigl[
    \sum_{\alpha=1}^K
    \nabla_{(a,w)}\widehat\Psi(x;(a,w),q_t)^\top
    \nabla_1\ell(H(x;\rho_t,q_t),y)
  \Bigr]\\
  &=
  K\,B_{\mathrm{single}}((a,w);\rho_t,q_t).
\end{align*}
Therefore the transport equation \eqref{eq:pde} reduces to
\eqref{eq:diag-single-pde}.

For the boundary component, \cref{eq:boundary-drift} yields
\[
  \dot q_t
  =
  C(\rho_t,q_t)
  =
  C_{\mathrm{single}}(\rho_t,q_t),
\]
which is exactly \eqref{eq:diag-single-ode}. This proves the claim.
\end{proof}

\subsection{Collapsed Boundary-Only Effective Kernel}

\begin{proof}[Proof of \cref{cor:boundary-single-kernel}]
Let $\Phi_t$ be the characteristic flow generated by
\[
  \dot\Phi_t(a,w)=K\,B_{\mathrm{single}}(\Phi_t(a,w);\rho_t,q_t),
  \qquad
  \rho_t=(\Phi_t)_\#\rho_0.
\]
Since
\[
  f(x,t)=H(x;\rho_t,q_t)
  =
  \int_{\R^{d_y}\times\R^{d_x}}
  \widehat\Psi(x;\Phi_t(a,w),q_t)\,\rho_0(\dd a,\dd w),
\]
the assumed regularity and the bounded first derivatives of $\widehat\Psi$
allow differentiation under the integral sign. Hence
  \begin{align*}
    \partial_t f(x,t)
    &=
    \int_{\R^{d_y}\times\R^{d_x}}
    \nabla_{(a,w)}\widehat\Psi(x;(a,w),q_t)^\top
    K\,B_{\mathrm{single}}((a,w);\rho_t,q_t)\,\rho_t(\dd a,\dd w)\\
    &\quad
    +
    \nabla_q H(x;\rho_t,q_t)^\top \dot q_t.
  \end{align*}
Substituting \cref{eq:single-bdry-B,eq:diag-single-ode} gives
\begin{align*}
  \partial_t f(x,t)
  &=
  -\E_{(x',y)\sim\mathsf P}
  \Bigl[
    K
    \int_{\R^{d_y}\times\R^{d_x}}
    \nabla_{(a,w)}\widehat\Psi(x;(a,w),q_t)^\top
    \nabla_{(a,w)}\widehat\Psi(x';(a,w),q_t)\,
    \rho_t(\dd a,\dd w)\,
    \nabla_1\ell(f(x',t),y)
  \Bigr]\\
  &\quad
  -
  \E_{(x',y)\sim\mathsf P}
  \Bigl[
    \nabla_q H(x;\rho_t,q_t)^\top
    \Bigl(
      \int_{\R^{d_y}\times\R^{d_x}}
      \nabla_q\widehat\Psi(x';(a,w),q_t)\,\rho_t(\dd a,\dd w)
    \Bigr)
    \nabla_1\ell(f(x',t),y)
  \Bigr].
\end{align*}
Because $\widehat\Psi$ has bounded first derivatives, dominated convergence also
gives
\[
  \nabla_q H(x;\rho_t,q_t)
  =
  \int_{\R^{d_y}\times\R^{d_x}}
  \nabla_q\widehat\Psi(x;(a,w),q_t)\,\rho_t(\dd a,\dd w).
\]
Substituting this identity into the second term yields
\cref{eq:collapsed-bdry-output,eq:collapsed-bdry-keff,eq:collapsed-bdry-kneu,eq:collapsed-bdry-kbdry}.

For comparison, the ordinary single-model mean-field gradient flow associated
with $\mathcal R_{\mathrm{single}}$ would replace the shared characteristic
equation
\[
  \dot\Phi_t(a,w)=K\,B_{\mathrm{single}}(\Phi_t(a,w);\rho_t,q_t)
\]
by
\[
  \dot\Phi_t(a,w)=B_{\mathrm{single}}(\Phi_t(a,w);\rho_t,q_t),
\]
equivalently replacing \eqref{eq:diag-single-pde} by the same transport
equation without the factor $K$, while keeping
\(
  \dot q_t=C_{\mathrm{single}}(\rho_t,q_t)
\).
Repeating the same calculation would therefore produce
\(
  \Gamma_{\mathrm{neu}}^{\mathrm{bdry}}+\Gamma_{\mathrm{bdry}}^{\mathrm{bdry}}
\)
instead of
\(
  K\,\Gamma_{\mathrm{neu}}^{\mathrm{bdry}}+\Gamma_{\mathrm{bdry}}^{\mathrm{bdry}}
\),
which proves the final comparison statement.
\end{proof}

\subsection{Local Covariance Expansion}

\begin{proof}
Write
\[
  q_\star:=(\mathbf 1,\mathbf 1),
  \qquad
  \eta:=(\xi,\zeta)\in\R^{d_x+d_y}.
\]
Since $H_t$ is $C^2$ on $\mathcal N$ with bounded second derivative, the
Hilbert-space Taylor formula at $q_\star$ gives
\[
  Y_t^\varepsilon
  =
  H_t(q_\star)
  +
  \varepsilon\,D H_t(q_\star)[\eta]
  +
  R_t^\varepsilon,
\]
with the quadratic remainder bound
\[
  \|R_t^\varepsilon\|_{L^2(\mathsf P_X)}
  \le
  \frac12
  \sup_{q\in\mathcal N}\|D^2 H_t(q)\|\,
  \varepsilon^2\|\eta\|^2
  \qquad
  \text{a.s.}
\]
for all sufficiently small $\varepsilon$. Because $\eta$ is essentially bounded,
\[
  \E\bigl[\|R_t^\varepsilon\|_{L^2(\mathsf P_X)}^2\bigr]
  =
  O(\varepsilon^4).
\]
Moreover,
\[
  H_t(q_\star)=g_t,
  \qquad
  D H_t(q_\star)[\eta]=A_t\xi+B_t\zeta,
\]
which proves \eqref{eq:local-linearization}.

Set
\[
  Z_t:=A_t\xi+B_t\zeta.
\]
Because $\xi$ and $\zeta$ are centered,
\[
  \E[Z_t]=0.
\]
Therefore
\[
  Y_t^\varepsilon-g_t=\varepsilon Z_t+R_t^\varepsilon.
\]
Expanding the covariance operator around this decomposition gives
\begin{align*}
  \operatorname{Cov}(Y_t^\varepsilon)
  &=
  \E\Bigl[
    \bigl(\varepsilon Z_t+R_t^\varepsilon-\E[R_t^\varepsilon]\bigr)
    \otimes
    \bigl(\varepsilon Z_t+R_t^\varepsilon-\E[R_t^\varepsilon]\bigr)
  \Bigr]\\
  &=
  \varepsilon^2\E[Z_t\otimes Z_t]
  +
  \varepsilon\,\E\bigl[Z_t\otimes(R_t^\varepsilon-\E[R_t^\varepsilon])\bigr]\\
  &\quad+
  \varepsilon\,\E\bigl[(R_t^\varepsilon-\E[R_t^\varepsilon])\otimes Z_t\bigr]
  +
  \operatorname{Cov}(R_t^\varepsilon).
\end{align*}
The mixed terms are $o(\varepsilon^2)$ in operator norm by Cauchy--Schwarz,
since $Z_t\in L^2$ and
\(
  \|R_t^\varepsilon\|_{L^2(\Omega;L^2(\mathsf P_X))}
  =
  O(\varepsilon^2)
\).
The last term is $O(\varepsilon^4)$ in operator norm. Hence
\[
  \operatorname{Cov}(Y_t^\varepsilon)
  =
  \varepsilon^2\E[Z_t\otimes Z_t]
  +o(\varepsilon^2).
\]

It remains to identify $\E[Z_t\otimes Z_t]$. Expanding $Z_t$ gives
\[
  \E[Z_t\otimes Z_t]
  =
  A_t\Sigma_u A_t^*
  +
  B_t\Sigma_v B_t^*
  +
  A_t\Sigma_{uv} B_t^*
  +
  B_t\Sigma_{uv}^\top A_t^*,
\]
which is \eqref{eq:local-cov-operator}. This proves
\eqref{eq:local-cov-op}.

If the law of $Y_t^\varepsilon$ were Dirac for infinitely many
$\varepsilon\downarrow 0$, then its covariance operator would vanish for those
$\varepsilon$. But \eqref{eq:local-cov-op} shows that this is impossible when
$\mathcal K_t\neq 0$ and $\varepsilon$ is sufficiently small, proving the
non-Dirac conclusion.

For the concrete boundary-only model, differentiate \eqref{eq:H-bdry} at
$(u,v)=(\mathbf 1,\mathbf 1)$. The derivative in the input-side direction
$q\in\R^{d_x}$ is
\[
  (A_t q)(x)
  =
  \int
  a\,\sigma'\!\bigl(w^\top x\bigr)\,
  w^\top(q\odot x)\,\rho_t(\dd a,\dd w)
  =
  J_t(x)q,
\]
for the Jacobian matrix $J_t(x)=D_u H_t(\mathbf 1,\mathbf 1)(x)$. The
derivative in the output-side direction $r\in\R^{d_y}$ is
\[
  (B_t r)(x)
  =
  r\odot
  \int a\,\sigma\!\bigl(w^\top x\bigr)\,\rho_t(\dd a,\dd w)
  =
  \operatorname{Diag}(g_t(x))\,r.
\]
Substituting these formulas into \eqref{eq:local-cov-operator} yields the kernel
representation \eqref{eq:local-cov-kernel}.
\end{proof}

\subsection{Local Disagreement Trace}

\begin{proof}
Let $Y$ be a square-integrable random variable taking values in the Hilbert
space $L^2(\mathsf P_X;\R^{d_y})$, and let $Y'$ be an iid copy. The standard
Hilbert-space identity gives
\[
  \E\bigl[\|Y-Y'\|_{L^2(\mathsf P_X)}^2\bigr]
  =
  2\,\operatorname{Tr}(\operatorname{Cov}(Y)).
\]
Applying this with $Y=Y_t^\varepsilon$ and using
\eqref{eq:local-cov-op} proves \eqref{eq:local-disagreement-trace}.

Since $A_t$ and $B_t$ have finite-dimensional domains, $\mathcal K_t$ is a
finite-rank operator, hence trace class. By cyclicity of the trace,
\begin{align*}
  \operatorname{Tr}(\mathcal K_t)
  &=
  \operatorname{Tr}(A_t\Sigma_u A_t^*)
  +
  \operatorname{Tr}(B_t\Sigma_v B_t^*)
  +
  \operatorname{Tr}(A_t\Sigma_{uv} B_t^*)
  +
  \operatorname{Tr}(B_t\Sigma_{uv}^\top A_t^*)\\
  &=
  \operatorname{Tr}(A_t^*A_t\Sigma_u)
  +
  \operatorname{Tr}(B_t^*B_t\Sigma_v)
  +
  \operatorname{Tr}(A_t^*B_t\Sigma_{uv})
  +
  \operatorname{Tr}(B_t^*A_t\Sigma_{uv}^\top),
\end{align*}
which is exactly \eqref{eq:local-design-matrix}.

Finally, if $\Sigma$ ranges over positive-semidefinite matrices with fixed trace
budget, then the spectral theorem shows that
\(
  \operatorname{Tr}(M_t\Sigma)
\)
is maximized by placing the covariance mass on the top eigendirections of the
symmetric matrix $M_t$.
\end{proof}

\end{document}
