\subsection{Linearized Symmetry Breaking: Consensus and Contrast Modes}

Proposition~\ref{prop:paper-collapse-baseline} identifies the correct deterministic reference trajectory on the permutation-symmetric diagonal. The next question is how small asymmetries evolve around that diagonal, and which ones generate useful disagreement rather than merely perturbing the common predictor.

Write the collapsed trajectory as
\[
(\rho_t^\star,q_t^\star)
\qquad
\text{with}
\qquad
q_t^\star=(u_t^\star,v_t^\star),
\]
so that
\[
\dot\rho_t^\star = K\,\mathcal B(\rho_t^\star,q_t^\star),
\qquad
\dot q_t^\star = \mathcal C(\rho_t^\star,q_t^\star),
\]
where $\mathcal C:=(\mathcal C_u,\mathcal C_v)$. Assume for this subsection that $\mathcal B$, $\mathcal C$, and the predictor map $H$ from \eqref{eq:member-state-predictor} are continuously differentiable along this path. Denote the corresponding derivatives by
\[
B_\rho(t):=D_\rho\mathcal B(\rho_t^\star,q_t^\star),
\quad
B_q(t):=D_q\mathcal B(\rho_t^\star,q_t^\star),
\]
\[
C_\rho(t):=D_\rho\mathcal C(\rho_t^\star,q_t^\star),
\quad
C_q(t):=D_q\mathcal C(\rho_t^\star,q_t^\star),
\]
and
\[
J_\rho(t,x):=D_\rho H(x;\rho_t^\star,q_t^\star),
\qquad
J_q(t,x):=D_q H(x;\rho_t^\star,q_t^\star).
\]

Consider a perturbation of the full $K$-member system of the form
\[
\rho_t = \rho_t^\star + \varepsilon \eta_t + o(\varepsilon),
\qquad
q_{\alpha,t} = q_t^\star + \varepsilon \zeta_{\alpha,t} + o(\varepsilon).
\]
Decompose the member perturbations into their empirical mean and centered part:
\[
\bar\zeta_t:=\frac{1}{K}\sum_{\alpha=1}^K \zeta_{\alpha,t},
\qquad
\xi_{\alpha,t}:=\zeta_{\alpha,t}-\bar\zeta_t,
\qquad
\sum_{\alpha=1}^K \xi_{\alpha,t}=0.
\]

\begin{theorem}[Consensus/contrast decomposition around the collapsed diagonal]
\label{thm:paper-consensus-contrast}
The linearized dynamics split into two invariant sectors:
\begin{align}
\dot\eta_t
&=
K\,B_\rho(t)\eta_t + K\,B_q(t)\bar\zeta_t,
\label{eq:paper-lin-rho}\\
\dot{\bar\zeta}_t
&=
C_\rho(t)\eta_t + C_q(t)\bar\zeta_t,
\label{eq:paper-lin-mean}\\
\dot\xi_{\alpha,t}
&=
C_q(t)\xi_{\alpha,t},
\qquad
\alpha=1,\dots,K.
\label{eq:paper-lin-contrast}
\end{align}
Moreover, the member predictors satisfy
\begin{align}
f_\alpha(x,t)-\bar f(x,t)
=
\varepsilon J_q(t,x)\xi_{\alpha,t} + o(\varepsilon),
\label{eq:paper-lin-disagreement}
\end{align}
whereas the ensemble mean satisfies
\begin{align}
\bar f(x,t)
=
H(x;\rho_t^\star,q_t^\star)
\,+\,
\varepsilon\bigl[J_\rho(t,x)\eta_t + J_q(t,x)\bar\zeta_t\bigr]
\,+\,
o(\varepsilon).
\label{eq:paper-lin-mean-predictor}
\end{align}
Hence only the centered contrast modes $\{\xi_{\alpha,t}\}_{\alpha=1}^K$ contribute to disagreement at first order, while the consensus mode $(\eta_t,\bar\zeta_t)$ controls the first-order shift of the common predictor.
\end{theorem}

\begin{proof}
Linearizing the full system
\[
\dot\rho_t = \sum_{\alpha=1}^K \mathcal B(\rho_t,q_{\alpha,t}),
\qquad
\dot q_{\alpha,t} = \mathcal C(\rho_t,q_{\alpha,t})
\]
around the diagonal trajectory $(\rho_t^\star,q_t^\star,\dots,q_t^\star)$ gives
\[
\dot\eta_t
=
\sum_{\alpha=1}^K \bigl(B_\rho(t)\eta_t + B_q(t)\zeta_{\alpha,t}\bigr)
=
K\,B_\rho(t)\eta_t + K\,B_q(t)\bar\zeta_t,
\]
and
\[
\dot\zeta_{\alpha,t}
=
C_\rho(t)\eta_t + C_q(t)\zeta_{\alpha,t}.
\]
Averaging the second display over $\alpha$ yields \eqref{eq:paper-lin-mean}, and subtracting the averaged equation from the $\alpha$th equation yields \eqref{eq:paper-lin-contrast}. Linearizing the predictor map gives
\[
f_\alpha(x,t)
=
H(x;\rho_t^\star,q_t^\star)
\,+\,
\varepsilon\bigl[J_\rho(t,x)\eta_t + J_q(t,x)\zeta_{\alpha,t}\bigr]
\,+\,
o(\varepsilon),
\]
from which \eqref{eq:paper-lin-disagreement} and \eqref{eq:paper-lin-mean-predictor} follow after writing $\zeta_{\alpha,t}=\bar\zeta_t+\xi_{\alpha,t}$ and averaging.
\end{proof}

Theorem~\ref{thm:paper-consensus-contrast} has a sharp design consequence. If one wants disagreement without perturbing the common predictor, then the perturbation budget should be placed entirely in the centered contrast sector.

Let $\Phi_C(t,s)$ denote the fundamental solution of \eqref{eq:paper-lin-contrast}, let
\[
\Sigma_0
:=
\frac{1}{K}\sum_{\alpha=1}^K \xi_{\alpha,0}\xi_{\alpha,0}^\top,
\]
and define the propagated contrast Gram matrix
\[
G_t(x)
:=
\Phi_C(t,0)^\top J_q(t,x)^\top J_q(t,x)\Phi_C(t,0).
\]

\begin{corollary}[Leading disagreement depends only on contrast covariance]
\label{cor:paper-contrast-cov}
If the initial perturbation is centered in the sense that
\[
\eta_0=0,
\qquad
\bar\zeta_0=0,
\]
then
\begin{equation}
\mathrm{Dis}(x,t)
=
\varepsilon^2 \operatorname{tr}\!\bigl(G_t(x)\Sigma_0\bigr)
\,+\,
o(\varepsilon^2).
\label{eq:paper-contrast-cov}
\end{equation}
Thus the leading disagreement depends on the initialization only through the empirical covariance of the contrast codes.
\end{corollary}

\begin{proof}
Under $\eta_0=0$ and $\bar\zeta_0=0$, the homogeneous system \eqref{eq:paper-lin-rho}--\eqref{eq:paper-lin-mean} implies $\eta_t\equiv 0$ and $\bar\zeta_t\equiv 0$. Hence
\[
\xi_{\alpha,t}=\Phi_C(t,0)\xi_{\alpha,0}
\]
and
\[
f_\alpha(x,t)-\bar f(x,t)
=
\varepsilon J_q(t,x)\Phi_C(t,0)\xi_{\alpha,0} + o(\varepsilon).
\]
Substituting this into the definition of $\mathrm{Dis}(x,t)$ yields \eqref{eq:paper-contrast-cov}.
\end{proof}

Equation~\eqref{eq:paper-contrast-cov} turns small-budget diversity design into a covariance-design problem. In particular, under the trace constraint
\[
\operatorname{tr}(\Sigma_0)\le \sigma^2,
\]
the optimal leading-order disagreement coefficient is $\sigma^2\lambda_{\max}(G_t(x))$ and is achieved by a rank-one antithetic code: half the heads are initialized with $+\sigma v_{\max}$ and half with $-\sigma v_{\max}$, where $v_{\max}$ is a top eigenvector of $G_t(x)$. This shows that, near the collapsed diagonal, iid isotropic randomness is generally suboptimal, and increasing $K$ by itself does not increase the best achievable first-order disagreement coefficient.

Furthermore, Theorem~\ref{thm:paper-consensus-contrast} implies that centered contrast perturbations do not move the ensemble mean predictor at order $\varepsilon$. Consequently, under the additional mild assumption that the solution map and the loss are $C^2$ in the perturbation scale, the ensemble risk of the mean predictor changes only at order $O(\varepsilon^2)$. So, to leading order, one can increase disagreement without paying a first-order price in the central predictor.
